### 2024-03-07 S. Bikulcius Thesis/analysis_report.Rmd
created analysis report rmd file

---
title: 'S. Bikulcius Thesis: Analysis'
author: "Micah E. Hirsch"
date: "2024-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

The purpose of this document is to report the results from the data analysis for S. Bikulcius' undergraduate thesis project. Analyses were conducted using R version 4.3.1.

```{r}

library(tidyverse)
library(rio)
library(glmmTMB)
library(janitor)
library(ggpubr)
library(patchwork)

```

# Participant Demographics

```{r pressure, echo=FALSE}

plot(pressure)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2024-03-08 S. Bikulcius Thesis/analysis_report.Rmd
small changes to rmd
listener descriptives

@@
 
+# Loading needed packages
+
 library(tidyverse)
@@
 library(patchwork)
+library(gt)
+library(gtsummary)
+library(ggridges)
+
+# Loading saved data
+
+## Pupil Dilation Data
+
+pupil <- rio::import("Data/cleaned_data.csv") |>
+  # Getting Analysis Window
+  dplyr::filter(timebins >= 500) |>
+  dplyr::group_by(subject, trial) |>
+  dplyr::filter(timebins <= max(timebins) - 2000) |>
+  dplyr::ungroup() |>
+  # Getting Peak Pupil Dilation Value
+  dplyr::group_by(subject, trial, speaker, code, rep_acc) |>
+  dplyr::summarize(peak_pupil = max(pupil.binned)) |>
+  dplyr::ungroup()
+
+## Cognitive Data
+
+cog <- rio::import("Data/cleaned_cog_data.csv")
+
+## Listener Demographics
+
+demo <- rio::import("Data/cleaned_listener_demo.csv") |>
+  dplyr::mutate(gender = factor(gender, levels = c("Man", "Woman", "Nonbinary", "Questioning", "Prefer not to answer")),
+                ethnicity = factor(ethnicity, levels = c("Hispanic/Latino(a/e)", "Not Hispanic/Latino(a/e)", "Prefer not to answer")),
+                race = factor(race, levels = c("white/Caucasian", "Black/African American", "Asian/Asian American",
+                                               "Native Hawaiian or Other Pacific Islander", "Native American or Alaska Native",
+                                               "Biracial or Multiracial", "Prefer not to answer", "Race not listed")),
+                native_lang = factor(native_lang, c("American English", "Not American English")))
+```
+
+# Listener Characteristics
+
+## Demographics
+
+The participant age range is from 18 to 39 years old.
+All participants reported they were fluent in English.
+However, two participants indicated that their native language is not American English.
+Their native languages were Spanish and Turkish respectively.
+
+```{r}
+
+demo_table <- demo |>
+  dplyr::select(age, gender, ethnicity, race, native_lang) |>
+  tbl_summary(type = list(age ~ "continuous",
+                          gender ~ "categorical",
+                          ethnicity ~ "categorical",
+                          race ~ "categorical",
+                          native_lang ~ "categorical"),
+              statistic = list(all_continuous() ~ "{mean} ({sd})",
+                               all_categorical() ~ "{n} ({p}%)"),
+              digits = list(everything() ~ c(2)),
+              label = list(age ~ "Age",
+                           gender ~ "Gender",
+                           ethnicity ~ "Ethnicity",
+                           race ~ "Race",
+                           native_lang ~ "Native Language")) |>
+  as_gt()
+
+demo_table
+
+demo_table |>
+  gt::gtsave("Tables/demo_table.html")
 
@@
 
-# Participant Demographics
+## Cognition
 
-```{r pressure, echo=FALSE}
+### Descriptives
 
-plot(pressure)
+```{r}
+
+cog_table <- cog |>
+  dplyr::select(-subject) |>
+  # Just using the age-corrected scores
+  dplyr::select(ends_with("_c")) |>
+  tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})"),
+              digits = list(everything() ~ c(2)),
+              label = list(list_sort_c ~ "Working Memory",
+                           flanker_c ~ "Inhibitory Control",
+                           card_sort_c ~ "Cognitive Flexibility",
+                           pattern_c ~ "Processing Speed")) |>
+  as_gt()
+
+cog_table
 
@@
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+### Distribution
+
+```{r}
+
+cog |>
+  dplyr::select(subject, ends_with("_c")) |>
+  tidyr::pivot_longer(cols = list_sort_c:pattern_c,
+                      names_to = "cog_measure",
+                      values_to = "score") |>
+  ggplot() +
+  aes(x = score,
+      y = cog_measure,
+      color = cog_measure,
+      fill = cog_measure) +
+  geom_density_ridges(jittered_points = T,
+                      position = position_points_jitter(width = 0.05, height = 0),
+                      point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.5) +
+  theme_classic()
+
+```

### 2024-03-08 S. Bikulcius Thesis/data_cleaning.R

@@ -15,7 +15,7 @@ library(janitor)
 
 # Setting Working Directory
 
-setwd("~/Documents/LE-Cognitive-Study/Data/Raw")
+setwd("~/Documents/LE-Cognitive-Study/Raw Data")
 
 # Loading in cognitive data
 
@@ -39,6 +39,6 @@ cog_data <- rio::import("2024-02-26 16.32.44 Assessment Scores.csv") |>
 
 # Exporting cleaned data for analysis
 
-setwd("~/Documents/LE-Cognitive-Study/Data")
+setwd("~/Documents/LE-Cognitive-Study/S. Bikulcius Thesis/Data")
 
 rio::export(cog_data, "cleaned_cog_data.csv")

### 2024-03-08 S. Bikulcius Thesis/analysis_report.Rmd
small edits

@@
 
+It appears that the listeners in this analysis had about average working memory and cognitive flexibility skills.
+Their inhibitory control scores were slightly below average and they tend to have higher processing speed scores (on the heigher end in the average range).
+The score distributions for working memory, cognitive flexibility, and inhibitory control roughly follow a normal distribution.
+The scores for processing speed appears to have more variation overall, and potentially has a bimodal distribution.
+
 ### Descriptives
@@
 ```
+
+```{r}
+
+rm(demo, demo_table)
+
+```
+
+# Speaker Characteristics
+
+```{r}
+
+```

### 2024-03-08 S. Bikulcius Thesis/analysis_report.Rmd
speaker descriptives

@@
   # Getting Peak Pupil Dilation Value
-  dplyr::group_by(subject, trial, speaker, code, rep_acc) |>
+  dplyr::group_by(subject, trial, speaker, code, rep_acc, target_number, correct_words) |>
   dplyr::summarize(peak_pupil = max(pupil.binned)) |>
@@
 
+cog_table |>
+  gt::gtsave("Tables/cog_table.html")
+
 ```
@@
 
+pupil <- pupil |>
+  # removing trials with missing rep accuracy
+  dplyr::mutate(rep_acc = ifelse(rep_acc == "", NA, rep_acc)) |>
+  dplyr::filter(!is.na(rep_acc)) |>
+  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
+                trial_c = trial - 6,
+                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")))
+
+speaker_table <- pupil |>
+  dplyr::select(subject, trial, speaker, correct_words, target_number, rep_acc) |>
+  dplyr::distinct() |>
+  dplyr::mutate(intel = (correct_words/target_number)*100) |>
+  dplyr::select(-c(subject, trial, correct_words, target_number)) |>
+  tbl_summary(by = speaker,
+              type = list(intel ~ "continuous",
+                          rep_acc ~ "categorical"),
+              statistic = list(all_continuous() ~ "{mean} ({sd})",
+                               all_categorical() ~ "{n} ({p}%)"),
+              digits = list(everything() ~ c(2)),
+              label = list(intel ~ "Intelligibility",
+                           rep_acc ~ "Repetition Accuracy")) |>
+  as_gt()
+
+speaker_table
+
+speaker_table |>
+  gt::gtsave("Tables/speaker_table.html")
+
 ```
+
+```{r}
+
+rm(speaker_table, cog_table)
+
+```

### 2024-03-13 S. Bikulcius Thesis/analysis_report.Rmd
pupil dilation models

@@
 library(ggridges)
+library(performance)
+library(sjPlot)
+library(ggcorrplot)
+library(lme4)
+library(lmerTest)
+library(optimx)
 
@@
   # Getting Peak Pupil Dilation Value
-  dplyr::group_by(subject, trial, speaker, code, rep_acc, target_number, correct_words) |>
+  dplyr::group_by(subject, trial, speaker, code, rep_acc, target_number, correct_words, effort_rating) |>
   dplyr::summarize(peak_pupil = max(pupil.binned)) |>
@@
                                                "Biracial or Multiracial", "Prefer not to answer", "Race not listed")),
-                native_lang = factor(native_lang, c("American English", "Not American English")))
+                native_lang = factor(native_lang, c("American English", "Not American English"))) |>
+  dplyr::rename(subject = id)
 ```
@@
 It appears that the listeners in this analysis had about average working memory and cognitive flexibility skills.
-Their inhibitory control scores were slightly below average and they tend to have higher processing speed scores (on the heigher end in the average range).
+Their inhibitory control scores were slightly below average and they tend to have higher processing speed scores (on the higher end in the average range).
 The score distributions for working memory, cognitive flexibility, and inhibitory control roughly follow a normal distribution.
@@
 
+### Correlations Among Cognitive Measures
+
+For our analysis, it is important to consider the correlations between working memory, inhibitory control, cognitive flexibility, and processing speed.
+The figure below shows a correlation matrix.
+The pearson correlation coefficients are in each block and the color of the block relates to how strong the correlations are.
+Values closer to 1 or -1 suggest a stronger positive or negative correlation between the two cognitive measures, respectively.
+
+The correlation matrix below shows a moderate positive correlation between cognitive flexibility and processing speed.
+There are also weak-to-moderate correlations between inhibitory control and cognitive flexibility and processing speed and working memory.
+This means the following:
+
+-   Listeners with higher processing speed scores also tend to have higher processing speed scores
+-   Listeners with higher inhibitory control scores also tend to have higher cognitive flexibility scores
+-   Listeners with higher processing speed scores also tend to have higher working memory scores
+
+```{r}
+
+cog |>
+  dplyr::select(ends_with("_c")) |>
+  dplyr::rename('Working Memory' = list_sort_c,
+                'Inhibitory Control' = flanker_c,
+                'Cognitive Flexibility' = card_sort_c,
+                'Processing Speed' = pattern_c) |>
+  cor(use = "pairwise.complete.obs") |>
+  ggcorrplot(
+    type = "lower",
+    outline.col = "white",
+    lab = TRUE,
+    insig = "blank",
+    legend.title = "Correlation"
+  )
+
+```
+
 # Speaker Characteristics
+
+In this study, we use speech samples from two speakers: a control speaker with neurotypical speech and a speaker with dysarthria secondary to Amyotropic Lateral Sclerosis (ALS).
+The mean speaker intelligibility is reported below.
+This was calculated using the listener responses.
+Although the speaker with ALS has a mean intelligibility of approximately 82%, about listeners still repeated approximately half of their trials incorrectly.
+When listening to the control speaker, there was a higher proportion of accurate responses.
 
@@
                 trial_c = trial - 6,
-                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")))
+                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate"))) |>
+  # recoding missing effort ratings (negative effort ratings indicate missing data)
+  dplyr::mutate(effort_rating = ifelse(effort_rating < 0, NA, effort_rating))
 
@@
 ```
+
+# Listening Effort
+
+## Descriptives
+
+### Listening Effort For Intelligible Phrases
+
+Mean peak pupil dilation values (arbitrary units) and mean effort ratings for accurately-recognized phrases for each speaker is reported in the table below.
+
+```{r}
+
+effort_table <- pupil |>
+  dplyr::filter(rep_acc == "accurate") |>
+  dplyr::select(speaker, peak_pupil, effort_rating) |>
+  tbl_summary(
+        by = speaker,
+        type = list(peak_pupil ~ "continuous",
+                    effort_rating ~ "continuous"),
+        statistic = list(all_continuous() ~ c("{mean} ({sd})")),
+        missing = "no",
+        digits = all_continuous() ~ 2,
+        label = list(peak_pupil ~ "Peak Pupil Dilation",
+                     effort_rating ~"Perceived Listening Effort")
+      ) |>
+  as_gt()
+
+effort_table
+
+effort_table |>
+  gt::gtsave("Tables/effort_table.html")
+
+```
+
+### Listening Effort: Comparing Intelligible and Non-Intelligible Phrases
+
+Mean peak pupil dilation values (arbitrary units) and mean effort ratings for when listener's accurately and inaccurately recognized the phrase for each speaker is reported below.
+
+```{r}
+
+effort_table2 <- pupil |>
+  dplyr::select(speaker, peak_pupil, effort_rating, rep_acc) |>
+  tbl_strata(
+    strata = rep_acc,
+    ~.x |>
+      tbl_summary(
+        by = speaker,
+        type = list(peak_pupil ~ "continuous",
+                    effort_rating ~ "continuous"),
+        statistic = list(all_continuous() ~ c("{mean} ({sd})")),
+        missing = "no",
+        digits = all_continuous() ~ 2,
+        label = list(peak_pupil ~ "Peak Pupil Dilation",
+                     effort_rating ~"Perceived Listening Effort")
+      )) |>
+  as_gt()
+
+effort_table2
+
+effort_table2 |>
+  gt::gtsave("Tables/effort_table_accuracy.html")
+
+```
+
+```{r}
+
+rm(effort_table, effort_table2)
+
+```
+
+## Pupil Dilation
+
+### Visualizations
+
+#### Overall Peak Pupil Dilation
+
+Figure showing peak pupil dilation for accurately recognized phrases for the control and ALS speakers.
+
+```{r}
+
+pupil_des <- pupil |>
+  dplyr::group_by(speaker, rep_acc) |>
+  dplyr::summarize(pupil = mean(peak_pupil, na.rm = T), 
+                   sd = sd(peak_pupil, na.rm = T),
+                   se = sd/sqrt(n())) |>
+  ungroup()
+
+pupil_des |>
+  dplyr::filter(rep_acc == "accurate") |>
+  ggplot() +
+   aes(x = speaker,
+       y = pupil,
+       group = speaker,
+       color = speaker,
+       fill = speaker) +
+  geom_bar(stat = "identity", alpha = 0.5) +
+  geom_errorbar(aes(ymin = pupil - se, ymax = pupil + se), width = 0.4) +
+  labs(x = "Speaker", y = "Peak Pupil Dilation (Arbitrary Units)") +
+  theme_bw() +
+  theme(aspect.ratio = 1)
+
+```
+
+Figure showing peak pupil dilation for accurately and inaccurately recognized phrases for the control and ALS speakers.
+
+```{r}
+
+pupil_des |>
+  ggplot() +
+   aes(x = rep_acc,
+       y = pupil,
+       group = speaker,
+       color = speaker,
+       fill = speaker) +
+  geom_bar(stat = "identity", alpha = 0.5, position = position_dodge()) +
+  geom_errorbar(aes(ymin = pupil - se, ymax = pupil + se), width = 0.4, position = position_dodge(.9)) +
+  labs(x = "Accuracy", y = "Peak Pupil Dilation (Arbitrary Units)") +
+  theme_bw() +
+  theme(aspect.ratio = 1)
+
+```
+
+#### Pupil Dilation Predicted by Listener Cognition
+
+Average Peak Pupil Dilation vs Cognitive Scores
+
+```{r}
+
+pupil1 <- dplyr::left_join(pupil, cog, by = "subject")
+
+pupil1 |>
+  dplyr::filter(rep_acc == "accurate") |>
+  dplyr::select(subject, trial, speaker, code, peak_pupil, ends_with("_c")) |>
+  dplyr::group_by(subject, speaker, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
+  dplyr::summarize(pupil = mean(peak_pupil)) |>
+  tidyr::pivot_longer(cols = list_sort_c:pattern_c,
+                     names_to = "cog_measure",
+                     values_to = "score") |>
+  ungroup() |>
+  ggplot() +
+  aes(x = score,
+      y = pupil,
+      color = speaker,
+      ) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = "lm") +
+  facet_wrap("cog_measure") +
+  theme_classic() +
+  theme(aspect.ratio = 1)
+
+```
+
+```{r}
+
+pupil1 |>
+  dplyr::select(subject, trial, speaker, code, rep_acc, peak_pupil, ends_with("_c")) |>
+  dplyr::group_by(subject, speaker, rep_acc, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
+  dplyr::summarize(pupil = mean(peak_pupil)) |>
+  tidyr::pivot_longer(cols = list_sort_c:pattern_c,
+                     names_to = "cog_measure",
+                     values_to = "score") |>
+  ungroup() |>
+  ggplot() +
+  aes(x = score,
+      y = pupil,
+      color = rep_acc,
+      ) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = "lm") +
+  facet_grid(cog_measure ~ speaker) +
+  theme_classic() +
+  theme(aspect.ratio = 1)
+
+ggsave("Figures/peak_pupil_cog_plot.png", plot = last_plot())
+
+```
+
+### Analysis
+
+Given our sample size and the primary question for this thesis project, we are just going to analyze the accurately recognized phrases for each speaker (i.e.
+100% intelligible phrases).
+To analyze the relationships between listener cognition and listening effort, we are using linear mixed effects models.
+
+#### Unconditional Model
+
+```{r}
+
+accurate_df <- pupil1 |>
+  dplyr::filter(rep_acc == "accurate")
+
+m0 <- glmmTMB(peak_pupil ~ 1 + (1|subject) + (1|code), data = accurate_df)
+summary(m0)
+performance::icc(m0)
+
+```
+
+#### Model 1
+
+```{r}
+
+m1 <- glmmTMB(peak_pupil ~ trial_c + (1|subject) + (1|code), data = accurate_df)
+summary(m1)
+performance::compare_performance(m0, m1)
+performance::test_performance(m0, m1)
+
+```
+
+#### Model 2
+
+```{r}
+
+m2 <- glmmTMB(peak_pupil ~ trial_c + speaker + (1|subject) + (1|code), data = accurate_df)
+summary(m2)
+performance::compare_performance(m1, m2)
+performance::test_performance(m1, m2)
+
+```
+
+#### Model 2b
+
+```{r}
+
+m2b <- glmmTMB(peak_pupil ~ trial_c + speaker + (1|subject), data = accurate_df)
+summary(m2b)
+performance::compare_performance(m2, m2b)
+performance::test_performance(m2, m2b)
+
+```
+
+#### Model 3
+
+```{r}
+
+m3 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c + (1|subject) + (1|code), data = accurate_df)
+summary(m3)
+performance::compare_performance(m2, m3)
+performance::test_performance(m2, m3)
+
+```
+
+#### Model 3b
+
+```{r}
+
+m3b <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c + (1|subject), data = accurate_df)
+summary(m3b)
+performance::compare_performance(m3, m3b)
+performance::test_performance(m3, m3b)
+
+```
+
+#### Comparing Model 3b with Model 2
+
+```{r}
+
+performance::compare_performance(m2, m3b)
+performance::test_performance(m2, m3b)
+
+```
+
+#### Model 4
+
+```{r}
+
+m4 <- glmmTMB(peak_pupil ~ trial_c + speaker + flanker_c + (1|subject), data = accurate_df)
+summary(m4)
+performance::compare_performance(m2b, m4)
+performance::test_performance(m2b, m4)
+
+```
+
+#### Model 5
+
+```{r}
+
+m5 <- glmmTMB(peak_pupil ~ trial_c + speaker + card_sort_c + (1|subject), data = accurate_df)
+summary(m5)
+performance::compare_performance(m2b, m5)
+performance::test_performance(m2b, m5)
+
+```
+
+#### Model 6
+
+```{r}
+
+m6 <- glmmTMB(peak_pupil ~ trial_c + speaker + pattern_c + (1|subject), data = accurate_df)
+summary(m6)
+performance::compare_performance(m2b, m6)
+performance::test_performance(m2b, m6)
+
+```
+
+#### Model 7
+
+```{r}
+
+m7 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + (1|subject), data = accurate_df)
+summary(m7)
+performance::compare_performance(m2b, m7)
+performance::test_performance(m2b, m7)
+
+```
+
+#### Model 8
+
+```{r}
+
+m8 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + flanker_c*speaker + (1|subject), data = accurate_df)
+summary(m8)
+performance::compare_performance(m7, m8)
+performance::test_performance(m7, m8)
+
+```
+
+#### Model 9
+
+```{r}
+
+m9 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + card_sort_c*speaker + (1|subject), data = accurate_df)
+summary(m9)
+performance::compare_performance(m7, m9)
+performance::test_performance(m7, m9)
+
+```
+
+#### Model 10
+
+```{r}
+
+m10 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + pattern_c*speaker + (1|subject), data = accurate_df)
+summary(m10)
+performance::compare_performance(m7, m10)
+performance::test_performance(m7, m10)
+
+```
+
+#### Final Model
+
+```{r}
+
+summary(m7)
+
+```

### 2024-03-13 S. Bikulcius Thesis/data_cleaning.R

@@ -30,10 +30,10 @@ cog_data <- rio::import("2024-02-26 16.32.44 Assessment Scores.csv") |>
   tidyr::pivot_longer(cols = uncorrected_standard_score:age_corrected_standard_score,
                       names_to = "type",
                       values_to = "score") |>
-  dplyr::mutate(measure = ifelse(grepl("uncorrected", type, ignore.case = T), 
+  dplyr::mutate(measure1 = ifelse(grepl("uncorrected", type, ignore.case = T), 
                                   paste(measure, "u", sep = "_"), paste(measure, "c", sep = "_"))) |>
-  dplyr::select(-type) |>
-  tidyr::pivot_wider(names_from = measure,
+  dplyr::select(-c(type, measure)) |>
+  tidyr::pivot_wider(names_from = measure1,
                      values_from = score) |>
   dplyr::rename(subject = pin)
 

### 2024-03-13 S. Bikulcius Thesis/analysis_report.Rmd
formatting pupil dilation results

@@
 
-```{r}
+```{r, echo = F, message = F, warning = F}
 
@@
 
-```{r}
+```{r, warning = F, message = F}
 
@@
   dplyr::select(subject, ends_with("_c")) |>
-  tidyr::pivot_longer(cols = list_sort_c:pattern_c,
+  dplyr::rename("Working Memory" = list_sort_c,
+                "Inhibitory Control" = flanker_c,
+                "Cognitive Flexibility" = card_sort_c,
+                "Processing Speed" = pattern_c) |>
+  tidyr::pivot_longer(cols = 'Working Memory':'Processing Speed',
                       names_to = "cog_measure",
@@
                       point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.5) +
+  labs(x = "Score", y = "", color = "Cognitive Measure", fill = "Cognitive Measure") +
   theme_classic()
@@
 
-```{r}
+```{r, echo = F, message = F, warning = F}
 
-rm(demo, demo_table)
+rm(demo, demo_table, cog_table)
 
@@
 
-```{r}
+```{r, echo = F, message = F, warning = F}
 
-rm(speaker_table, cog_table)
+rm(speaker_table)
 
@@
 
-Mean peak pupil dilation values (arbitrary units) and mean effort ratings for accurately-recognized phrases for each speaker is reported in the table below.
+Mean peak pupil dilation values (arbitrary units) and mean effort ratings for accurately-recognized phrases (i.e. 100% intelligible) for each speaker is reported in the table below.
 
-```{r}
+```{r, warning = F, message = F}
 
@@
 
+##### Intelligible Phrases Only
+
 Figure showing peak pupil dilation for accurately recognized phrases for the control and ALS speakers.
 
-```{r}
+```{r, warning = F, message = F}
 
@@
   labs(x = "Speaker", y = "Peak Pupil Dilation (Arbitrary Units)") +
-  theme_bw() +
+  theme_classic() +
   theme(aspect.ratio = 1)
@@
 
+##### Comparing Intelligible vs. Non-Intelligible Phrases
+
 Figure showing peak pupil dilation for accurately and inaccurately recognized phrases for the control and ALS speakers.
 
-```{r}
+```{r, warning = F, message = F}
 
@@
   labs(x = "Accuracy", y = "Peak Pupil Dilation (Arbitrary Units)") +
-  theme_bw() +
+  theme_classic() +
   theme(aspect.ratio = 1)
@@
 
-Average Peak Pupil Dilation vs Cognitive Scores
+The figure below shows the relationship between each of the cognitive measures and peak pupil dilation for accurately recognized phrases.
+Each panel is a separate cognitive measure.
+We can see that there appears to be a positive relationship between peak pupil dilation and cognitive flexibility and inhibitory control for each speaker.
+There also appears to be a positive relationship between peak pupil dilation and working memory and processing speed for the ALS speaker.
+However, this relationship appears weaker when listeners were listening to the control speaker.
 
-```{r}
+```{r, warning = F, message = F}
 
@@
   dplyr::summarize(pupil = mean(peak_pupil)) |>
-  tidyr::pivot_longer(cols = list_sort_c:pattern_c,
+  dplyr::rename('Working Memory' = list_sort_c,
+                'Inhibitory Control' = flanker_c,
+                'Cognitive Flexibility' = card_sort_c,
+                'Processing Speed' = pattern_c) |>
+  tidyr::pivot_longer(cols = "Working Memory":"Processing Speed",
                      names_to = "cog_measure",
@@
   facet_wrap("cog_measure") +
+  labs(x = "Score", y = "Peak Pupil Dilation (Arbitrary Units)") +
   theme_classic() +
@@
 
-```{r}
+This figure is showing the same relationships between listener cognition and peak pupil dilation for each speaker.
+However, we are also now looking at inaccurate vs accurate phrase repetitions (i.e.
+intelligible vs unintelligible trials).
+
+```{r, warning = F, message = F}
 
@@
       y = pupil,
-      color = rep_acc,
+      color = speaker,
       ) +
@@
   geom_smooth(method = "lm") +
-  facet_grid(cog_measure ~ speaker) +
+  facet_grid(rep_acc ~ cog_measure) +
+  labs(x = "Score", y = "Peak Pupil Dilation (Arbitrary Units)") +
   theme_classic() +
@@
 To analyze the relationships between listener cognition and listening effort, we are using linear mixed effects models.
+In a linear mixed effects model, we can add fixed effects (i.e.
+our predictors that we are interested in modeling; speaker, working memory, etc) and random effects (i.e.
+variables that we aren't controlling bur we are accounting for; each individual listener, the phrases).
+We took a model building approach to this analysis.
+We start with the simplest model with no fixed or random effects (the unconditional model), and then we gradually add all our random and fixed effects to test whether adding them improves the model fit.
+We will eventually find the best fitting model which tells us which fixed effects are important for predicting peak pupil dilation.
+
+The test we use to compare models is called the Likelihood Ratio Test (LRT).
 
@@
 
-```{r}
+We start with an unconditional model because we need to calculate a value called the intraclass correlation coefficient.
+The ICC is .304 which tells us that a large portion of the variation in our pupil dilation data is due to the listener and phrases.
+
+```{r, warning = F, message = F}
 
@@
 m0 <- glmmTMB(peak_pupil ~ 1 + (1|subject) + (1|code), data = accurate_df)
-summary(m0)
+sjPlot::tab_model(m0, pred.labels = c("Intercept"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::icc(m0)
@@
 m1 <- glmmTMB(peak_pupil ~ trial_c + (1|subject) + (1|code), data = accurate_df)
-summary(m1)
+sjPlot::tab_model(m1, pred.labels = c("Intercept", "Trial Order"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::compare_performance(m0, m1)
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 m2 <- glmmTMB(peak_pupil ~ trial_c + speaker + (1|subject) + (1|code), data = accurate_df)
-summary(m2)
+sjPlot::tab_model(m2, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::compare_performance(m1, m2)
@@
 
-```{r}
+When I added the next predictor for model 3, I ran into convergence issues.
+After trying to use a different optimizer, I still had convergence issues, so I decided to take the random effect for phrase out.
+The next series of models (anything with b at the end) is to determine whether the model fit changes were primarily due to changes in the random effect structure or fixed effects.
+
+```{r, warning = F, message = F}
 
 m2b <- glmmTMB(peak_pupil ~ trial_c + speaker + (1|subject), data = accurate_df)
-summary(m2b)
+sjPlot::tab_model(m2b)
 performance::compare_performance(m2, m2b)
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 m3 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c + (1|subject) + (1|code), data = accurate_df)
-summary(m3)
+sjPlot::tab_model(m3)
 performance::compare_performance(m2, m3)
@@
 
-```{r}
+```{r, message = F, warning = F}
 
 m3b <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c + (1|subject), data = accurate_df)
-summary(m3b)
+sjPlot::tab_model(m3b)
 performance::compare_performance(m3, m3b)
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 m4 <- glmmTMB(peak_pupil ~ trial_c + speaker + flanker_c + (1|subject), data = accurate_df)
-summary(m4)
+sjPlot::tab_model(m4, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Inhibitory Control"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::compare_performance(m2b, m4)
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 m5 <- glmmTMB(peak_pupil ~ trial_c + speaker + card_sort_c + (1|subject), data = accurate_df)
-summary(m5)
+sjPlot::tab_model(m5, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Cognitive Flexibility"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::compare_performance(m2b, m5)
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 m6 <- glmmTMB(peak_pupil ~ trial_c + speaker + pattern_c + (1|subject), data = accurate_df)
-summary(m6)
+sjPlot::tab_model(m6, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Processing Speed"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::compare_performance(m2b, m6)
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 m7 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + (1|subject), data = accurate_df)
-summary(m7)
+sjPlot::tab_model(m7, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", 
+                                      "Working Memory * Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::compare_performance(m2b, m7)
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 m8 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + flanker_c*speaker + (1|subject), data = accurate_df)
-summary(m8)
+sjPlot::tab_model(m8, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", "Inhibitory Control",
+                                      "Working Memory * Speaker [ALS]", "Inhibitory Control * Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::compare_performance(m7, m8)
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 m9 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + card_sort_c*speaker + (1|subject), data = accurate_df)
-summary(m9)
+sjPlot::tab_model(m9, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", "Cognitive Flexibility",
+                                      "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::compare_performance(m7, m9)
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 m10 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + pattern_c*speaker + (1|subject), data = accurate_df)
-summary(m10)
+sjPlot::tab_model(m10, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", "Processing Speed",
+                                      "Working Memory * Speaker [ALS]", "Processing Speed * Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::compare_performance(m7, m10)
@@
 
-```{r}
+Model 7 was the best fitting model for our data. Therefore, model 7 is our final model. The model output is below. Here are the takeaways:
 
-summary(m7)
+- The fixed effects alone account for 11% of the variation in the peak pupil dilation.
+However, when including speaker as a random intercept, our model accounts for about 40% of the variation in pupil dilation.
+This is a good thing because you generally want your models to explain more variation in your outcome.
+- The fixed effect of trial order is significant indicating that peak pupil dilation decreased on average by 1.7 units across trials.
+
+```{r, warning = F, message = F}
+
+final <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c)*speaker + (1|subject), data = accurate_df)
+sjPlot::tab_model(final, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", 
+                                      "Working Memory * Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation")
 

### 2024-03-13 S. Bikulcius Thesis/analysis_report.Rmd
Figures for perceived listening effort

@@
 
+To simplify the model interpretation, we mean-centered the cognitive measure scores.
+
 The test we use to compare models is called the Likelihood Ratio Test (LRT).
@@
 
-m3 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c + (1|subject) + (1|code), data = accurate_df)
+m3 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c) + (1|subject) + (1|code), data = accurate_df)
 sjPlot::tab_model(m3)
@@
 
-m3b <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c + (1|subject), data = accurate_df)
+m3b <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c) + (1|subject), data = accurate_df)
 sjPlot::tab_model(m3b)
@@
 
-m4 <- glmmTMB(peak_pupil ~ trial_c + speaker + flanker_c + (1|subject), data = accurate_df)
+m4 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(flanker_c) + (1|subject), data = accurate_df)
 sjPlot::tab_model(m4, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Inhibitory Control"),
@@
 
-m5 <- glmmTMB(peak_pupil ~ trial_c + speaker + card_sort_c + (1|subject), data = accurate_df)
+m5 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(card_sort_c) + (1|subject), data = accurate_df)
 sjPlot::tab_model(m5, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Cognitive Flexibility"),
@@
 
-m6 <- glmmTMB(peak_pupil ~ trial_c + speaker + pattern_c + (1|subject), data = accurate_df)
+m6 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(pattern_c) + (1|subject), data = accurate_df)
 sjPlot::tab_model(m6, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Processing Speed"),
@@
 
-m7 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + (1|subject), data = accurate_df)
+m7 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c)*speaker + (1|subject), data = accurate_df)
 sjPlot::tab_model(m7, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", 
@@
 
-m8 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + flanker_c*speaker + (1|subject), data = accurate_df)
+m8 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c)*speaker + scale(flanker_c)*speaker + (1|subject), data = accurate_df)
 sjPlot::tab_model(m8, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", "Inhibitory Control",
@@
 
-m9 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + card_sort_c*speaker + (1|subject), data = accurate_df)
+m9 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c)*speaker + scale(card_sort_c)*speaker + (1|subject), data = accurate_df)
 sjPlot::tab_model(m9, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", "Cognitive Flexibility",
@@
 
-m10 <- glmmTMB(peak_pupil ~ trial_c + speaker + list_sort_c*speaker + pattern_c*speaker + (1|subject), data = accurate_df)
+m10 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c)*speaker + scale(pattern_c)*speaker + (1|subject), data = accurate_df)
 sjPlot::tab_model(m10, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", "Processing Speed",
@@
 This is a good thing because you generally want your models to explain more variation in your outcome.
-- The fixed effect of trial order is significant indicating that peak pupil dilation decreased on average by 1.7 units across trials.
+- The fixed effect of trial order is significant indicating that peak pupil dilation decreased on average by 1.7 units across trials, controlling for speaker and working memory.
+- The fixed effect for speaker is significant indicating that listeners' peak pupil dilation was higher when listening to the speaker with ALS compared to the control speaker, controlling for trial order and working memory.
+- The interaction between working memory and speaker is significant indicating that listeners with greater working memory on average had larger pupil dilation when listening to the speaker with ALS. 
+- The fixed effect of working memory is not significant, indicating that working memory was not predictive of peak pupil dilation when listening to the control speaker.
 
@@
 
-final <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c)*speaker + (1|subject), data = accurate_df)
-sjPlot::tab_model(final, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", 
+sjPlot::tab_model(m7, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", 
                                       "Working Memory * Speaker [ALS]"),
-                  dv.labels = "Peak Pupil Dilation")
+                  dv.labels = "Peak Pupil Dilation",
+                  file = "Tables/Peak_pupil_dilation_model.html")
 
 ```
+
+```{r, echo = F, warning = F, message = F}
+
+rm(m0, m1, m2, m2b, m3, m3b, m4, m5, m6, m7, m8, m9, m10, pupil_des)
+
+```
+
+## Perceived Listening Effort Ratings
+
+### Visualizations
+
+#### Overall Perceived LE Ratings
+
+The perceived listening effort ratings for accurately perceived phrases is shown in the figure below. Listeners rated the speaker with ALS as more effortful than the control speaker.
+
+```{r, warning = F, message = F}
+
+pupil_des <- pupil |>
+  dplyr::group_by(speaker, rep_acc) |>
+  dplyr::summarize(effort = mean(effort_rating, na.rm = T), 
+                   sd = sd(effort_rating, na.rm = T),
+                   se = sd/sqrt(n())) |>
+  ungroup()
+
+pupil_des |>
+  dplyr::filter(rep_acc == "accurate") |>
+  ggplot() +
+   aes(x = speaker,
+       y = effort,
+       group = speaker,
+       color = speaker,
+       fill = speaker) +
+  geom_bar(stat = "identity", alpha = 0.5) +
+  geom_errorbar(aes(ymin = effort - se, ymax = effort + se), width = 0.4) +
+  labs(x = "Speaker", y = "Perceived Listening Effort Rating") +
+  theme_classic() +
+  theme(aspect.ratio = 1)
+
+```
+
+The figure below shows perceived listening effort ratings for the two speakers for when the listener accurately perceived the phrase vs. when they inaccurately perceived the phrase.
+
+```{r, warning = F, message = F}
+
+pupil_des |>
+  ggplot() +
+   aes(x = rep_acc,
+       y = effort,
+       group = speaker,
+       color = speaker,
+       fill = speaker) +
+  geom_bar(stat = "identity", alpha = 0.5, position = position_dodge()) +
+  geom_errorbar(aes(ymin = effort - se, ymax = effort + se), width = 0.4, position = position_dodge(.9)) +
+  labs(x = "Accuracy", y = "Perceived Listening Effort Rating") +
+  theme_classic() +
+  theme(aspect.ratio = 1)
+
+```
+
+#### Perceived LE Ratings Predicted by Cognitive Measures
+
+```{r, warning = F, message = F}
+
+pupil1 |>
+  dplyr::filter(rep_acc == "accurate") |>
+  dplyr::select(subject, trial, speaker, code, effort_rating, ends_with("_c")) |>
+  dplyr::group_by(subject, speaker, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
+  dplyr::summarize(effort = mean(effort_rating)) |>
+  dplyr::rename('Working Memory' = list_sort_c,
+                'Inhibitory Control' = flanker_c,
+                'Cognitive Flexibility' = card_sort_c,
+                'Processing Speed' = pattern_c) |>
+  tidyr::pivot_longer(cols = "Working Memory":"Processing Speed",
+                     names_to = "cog_measure",
+                     values_to = "score") |>
+  ungroup() |>
+  ggplot() +
+  aes(x = score,
+      y = effort,
+      color = speaker,
+      ) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = "lm") +
+  facet_wrap("cog_measure") +
+  labs(x = "Score", y = "Perceived Listening Effort Rating") +
+  theme_classic() +
+  theme(aspect.ratio = 1)
+
+```
+
+```{r, warning = F, message = F}
+
+pupil1 |>
+  dplyr::select(subject, trial, speaker, code, rep_acc, effort_rating, ends_with("_c")) |>
+  dplyr::group_by(subject, speaker, rep_acc, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
+  dplyr::summarize(effort = mean(effort_rating)) |>
+  dplyr::rename('Working Memory' = list_sort_c,
+                'Inhibitory Control' = flanker_c,
+                'Cognitive Flexibility' = card_sort_c,
+                'Processing Speed' = pattern_c) |>
+  tidyr::pivot_longer(cols = "Working Memory":"Processing Speed",
+                     names_to = "cog_measure",
+                     values_to = "score") |>
+  ungroup() |>
+  ggplot() +
+  aes(x = score,
+      y = effort,
+      color = speaker,
+      ) +
+  geom_point(alpha = 0.3) +
+  geom_smooth(method = "lm") +
+  facet_grid(rep_acc ~ cog_measure) +
+  labs(x = "Score", y = "Perceived Listening Effort Rating") +
+  theme_classic() +
+  theme(aspect.ratio = 1)
+
+```
+
+### Analysis
+
+#### Unconditional Model
+
+```{r}
+
+```

### 2024-03-14 S. Bikulcius Thesis/analysis_report.Rmd
perceived listening effort analysis

@@
 
-m3 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c) + (1|subject) + (1|code), data = accurate_df)
+m3 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c, scale = F) + (1|subject) + (1|code), data = accurate_df)
 sjPlot::tab_model(m3)
@@
 
-m3b <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c) + (1|subject), data = accurate_df)
+m3b <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c, scale = F) + (1|subject), data = accurate_df)
 sjPlot::tab_model(m3b)
@@
 
-#### Comparing Model 3b with Model 2
+#### Comparing Model 3b with Model 2b
 
@@
 
-performance::compare_performance(m2, m3b)
-performance::test_performance(m2, m3b)
+performance::compare_performance(m2b, m3b)
+performance::test_performance(m2b, m3b)
 
@@
 
-m4 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(flanker_c) + (1|subject), data = accurate_df)
+m4 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(flanker_c, scale = F) + (1|subject), data = accurate_df)
 sjPlot::tab_model(m4, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Inhibitory Control"),
@@
 
-m5 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(card_sort_c) + (1|subject), data = accurate_df)
+m5 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(card_sort_c, scale = F) + (1|subject), data = accurate_df)
 sjPlot::tab_model(m5, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Cognitive Flexibility"),
@@
 
-m6 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(pattern_c) + (1|subject), data = accurate_df)
+m6 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(pattern_c, scale = F) + (1|subject), data = accurate_df)
 sjPlot::tab_model(m6, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Processing Speed"),
@@
 
-m7 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c)*speaker + (1|subject), data = accurate_df)
+m7 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c, scale = F)*speaker + (1|subject), data = accurate_df)
 sjPlot::tab_model(m7, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", 
@@
 
-m8 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c)*speaker + scale(flanker_c)*speaker + (1|subject), data = accurate_df)
+m8 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c, scale = F)*speaker + scale(flanker_c, scale = F)*speaker + (1|subject), data = accurate_df)
 sjPlot::tab_model(m8, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", "Inhibitory Control",
@@
 
-m9 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c)*speaker + scale(card_sort_c)*speaker + (1|subject), data = accurate_df)
+m9 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c, scale = F)*speaker + scale(card_sort_c, scale = F)*speaker + (1|subject), data = accurate_df)
 sjPlot::tab_model(m9, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", "Cognitive Flexibility",
@@
 
-m10 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c)*speaker + scale(pattern_c)*speaker + (1|subject), data = accurate_df)
+m10 <- glmmTMB(peak_pupil ~ trial_c + speaker + scale(list_sort_c, scale = F)*speaker + scale(pattern_c, scale = F)*speaker + (1|subject), data = accurate_df)
 sjPlot::tab_model(m10, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", "Processing Speed",
@@
 
+The figure below shows the relationships between each of the cognitive measures with perceived listening effort for accurately recognized phrases.
+The lines in the graph are for each speaker (the control and ALS speaker).
+We see an interesting trend here.
+Unlike with peak pupil dilation, we see more negative relationships between some of these cognitive measures and peak pupil dilation.
+It seems like processing speed and working memory has little to no relationship with perceived listening effort, especially for the ALS speaker.
+However, there may be a slight negative relationship between working memory and perceived listening effort for the control speaker.
+It seems like inhibitory control has a negative relationship with perceived listening effort ratings for both speakers.
+Cognitive flexibility appears to have a negative relationship with perceived listening effort for the ALS speaker, but positive for the control speaker.
+
 ```{r, warning = F, message = F}
@@
 ```
+
+The figure below shows the same relationship as the figure above.
+However, now we are comparing these relationships between when listeners accurately perceived the phrases vs when they did not.
+Overall, it seems like the trends are similar between accurately perceived phrases vs inaccurately perceived phrases.
 
@@
 
+Like with the pupil dilation data, we are using a linear mixed effects (LME) model to analyze the predictive relationships between the cognitive measures and perceived listening effort ratings.
+We are using the same model-building approach.
+
 #### Unconditional Model
@@
 
+m0 <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = accurate_df)
+summary(m0)
+sjPlot::tab_model(m0, pred.labels = c("Intercept"), dv.labels = "Perceived Listening Effort")
+performance::icc(m0)
+
 ```
+
+#### Model 1
+
+```{r}
+
+m1 <- glmmTMB(effort_rating ~ 1 + (trial_c|subject) + (trial_c|code), data = accurate_df)
+sjPlot::tab_model(m1, pred.labels = c("Intercept"), dv.labels = "Perceived Listening Effort")
+performance::compare_performance(m0, m1)
+performance::test_performance(m0, m1)
+
+```
+
+#### Model 2
+
+```{r}
+
+m2 <- glmmTMB(effort_rating ~ trial_c + (trial_c|subject) + (trial_c|code), data = accurate_df)
+sjPlot::tab_model(m2, pred.labels = c("Intercept", "Trial Order"), dv.labels = "Perceived Listening Effort")
+performance::compare_performance(m1, m2)
+performance::test_performance(m1, m2)
+
+```
+#### Model 3
+
+```{r}
+
+m3 <- glmmTMB(effort_rating ~ trial_c + speaker + (trial_c|subject) + (trial_c|code), data = accurate_df)
+sjPlot::tab_model(m3, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]"), dv.labels = "Perceived Listening Effort")
+performance::compare_performance(m2, m3)
+performance::test_performance(m2, m3)
+
+```
+
+#### Model 4
+
+Quick Note Here: Working Memory was a significant predictor.
+However, based on the likelihood ratio test, adding it to the model did not significantly improve model fit, which is why on subsequent steps, we are removing it.
+
+```{r}
+
+m4 <- glmmTMB(effort_rating ~ trial_c + speaker + scale(list_sort_c, scale = F) + (trial_c|subject) + (trial_c|code), data = accurate_df)
+sjPlot::tab_model(m4, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory"), dv.labels = "Perceived Listening Effort")
+performance::compare_performance(m3, m4)
+performance::test_performance(m3, m4)
+
+```
+#### Model 5
+
+```{r}
+
+m5 <- glmmTMB(effort_rating ~ trial_c + speaker + scale(flanker_c, scale = F) + (trial_c|subject) + (trial_c|code), data = accurate_df)
+sjPlot::tab_model(m5, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Inhibitory Control"), dv.labels = "Perceived Listening Effort")
+performance::compare_performance(m3, m5)
+performance::test_performance(m3, m5)
+
+```
+
+#### Model 6
+
+```{r}
+
+m6 <- glmmTMB(effort_rating ~ trial_c + speaker + scale(card_sort_c, scale = F) + (trial_c|subject) + (trial_c|code), data = accurate_df)
+sjPlot::tab_model(m6, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Cognitive Flexibility"), dv.labels = "Perceived Listening Effort")
+performance::compare_performance(m3, m6)
+performance::test_performance(m3, m6)
+
+```
+
+#### Model 7
+
+```{r}
+
+m7 <- glmmTMB(effort_rating ~ trial_c + speaker + scale(pattern_c, scale = F) + (trial_c|subject) + (trial_c|code), data = accurate_df)
+sjPlot::tab_model(m7, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Processing Speed"), dv.labels = "Perceived Listening Effort")
+performance::compare_performance(m3, m7)
+performance::test_performance(m3, m7)
+
+```
+#### Model 8
+
+```{r}
+
+m8 <- glmmTMB(effort_rating ~ trial_c + speaker + scale(list_sort_c, scale = F)*speaker + (trial_c|subject) + (trial_c|code), data = accurate_df)
+sjPlot::tab_model(m8, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", 
+                                      "Working Memory", "Working Memory * Speaker [ALS]"), 
+                  dv.labels = "Perceived Listening Effort")
+performance::compare_performance(m3, m8)
+performance::test_performance(m3, m8)
+
+```
+
+#### Model 9
+
+```{r}
+
+m9 <- glmmTMB(effort_rating ~ trial_c + speaker + scale(list_sort_c, scale = F)*speaker + 
+                scale(flanker_c, scale = F)*speaker +(trial_c|subject) + (trial_c|code), data = accurate_df)
+sjPlot::tab_model(m9, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", 
+                                      "Working Memory", "Inhibitory Control", "Working Memory * Speaker [ALS]",
+                                      "Inhibitory Control * Speaker [ALS]"), 
+                  dv.labels = "Perceived Listening Effort")
+performance::compare_performance(m8, m9)
+performance::test_performance(m8, m9)
+
+```
+
+#### Model 10
+
+```{r}
+
+m10 <- glmmTMB(effort_rating ~ trial_c + speaker + scale(list_sort_c, scale = F)*speaker + 
+                scale(card_sort_c, scale = F)*speaker + (trial_c|subject) + (trial_c|code), data = accurate_df)
+summary(m10)
+sjPlot::tab_model(m10, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", 
+                                      "Working Memory", "Cognitive Flexibility", "Working Memory * Speaker [ALS]",
+                                      "Cognitive Flexibility * Speaker [ALS]"), 
+                  dv.labels = "Perceived Listening Effort")
+performance::compare_performance(m8, m10)
+performance::test_performance(m8, m10)
+
+```
+
+#### Model 11
+
+```{r}
+
+m11 <- glmmTMB(effort_rating ~ trial_c + speaker + scale(list_sort_c, scale = F)*speaker + 
+                scale(card_sort_c, scale = F)*speaker + scale(pattern_c, scale = F)*speaker + (trial_c|subject) + (trial_c|code), data = accurate_df)
+summary(m11)
+sjPlot::tab_model(m11, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", 
+                                      "Working Memory", "Cognitive Flexibility", "Processing Speed",
+                                      "Working Memory * Speaker [ALS]",
+                                      "Cognitive Flexibility * Speaker [ALS]",
+                                      "Processing Speed * Speaker [ALS]"), 
+                  dv.labels = "Perceived Listening Effort")
+performance::compare_performance(m10, m11)
+performance::test_performance(m10, m11)
+
+```
+
+#### Final Model
+
+Based on the analysis, model 10 is the best fitting model to our data. The model takeaways are below.
+
+- Unlike with the pupil dilation data, trial order did not impact perceived listening effort ratings across trials.
+- The significant main effect for speaker indicates that perceived listening effort was higher for the ALS speaker compared to the control speaker, controlling for lsiteners' working memory and cognitive flexibility skills as well as trial order.
+- The main effect of working memory indicates that listeners with higher working memory scores on average gave lower perceived listening effort ratings for the control speaker, controlling for trial order and cognitive flexibility.
+- Although the main effect of cognitive flexibility is not significant, this result suggests that on average, listeners with higher cognitive flexibility rated perceived listening effort higher for the control speaker, controlling for trial order and working memory scores.
+This is a marginal finding, so it should be interpreted with caution.
+- The interaction between working memory and speaker is not significant, which indicates that the relationship between working memory and perceived listening effort when listening to the ALS speaker is not significantly different than the relationship for the control speaker, controlling for trial order and cognitive flexibility.
+However, in the model building steps, it should be noted that this interaction was significant before adding cognitive flexibility into the model.
+- The significant interaction between cognitive flexibility and speaker indicates that the relationship between cognitive flexibility and speaker is significantly different for the ALS speaker compared to the control speaker.
+In particular, the results suggest that there is a negative relationship between cognitive flexibility and perceived listening effort for the ALS speaker, controlling for trial order and listener working memory.
+In other words, listeners with higher cognitive flexibility scores had lower perceived listening effort ratings when they were listening to the speaker with ALS than those with lower cognitive flexibility scores.
+
+```{r}
+
+sjPlot::tab_model(m10, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", 
+                                      "Working Memory", "Cognitive Flexibility", "Working Memory * Speaker [ALS]",
+                                      "Cognitive Flexibility * Speaker [ALS]"), 
+                  dv.labels = "Perceived Listening Effort")
+
+```

### 2024-03-14 S. Bikulcius Thesis/analysis_report.Rmd

@@
 author: "Micah E. Hirsch"
-date: "2024-03-07"
-output: html_document
+date: "2024-03-14"
+output: 
+  html_document:
+    toc: true
+    toc_float: true
+    code_folding: hide
+
 ---
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 m0 <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = accurate_df)
-summary(m0)
 sjPlot::tab_model(m0, pred.labels = c("Intercept"), dv.labels = "Perceived Listening Effort")
@@
 Quick Note Here: Working Memory was a significant predictor.
-However, based on the likelihood ratio test, adding it to the model did not significantly improve model fit, which is why on subsequent steps, we are removing it.
+However, based on the likelihood ratio test, adding it to the model did not significantly improve model fit, which is why on subsequent steps, we are removing it for now.
+You'll see that it will eventually get added back in because of the significant interaction between working memory and speaker.
 
-```{r}
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+```{r, message = F, warning = F}
 
@@
                 scale(card_sort_c, scale = F)*speaker + (trial_c|subject) + (trial_c|code), data = accurate_df)
-summary(m10)
 sjPlot::tab_model(m10, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", 
@@
                 scale(card_sort_c, scale = F)*speaker + scale(pattern_c, scale = F)*speaker + (trial_c|subject) + (trial_c|code), data = accurate_df)
-summary(m11)
 sjPlot::tab_model(m11, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", 
@@
                                       "Cognitive Flexibility * Speaker [ALS]"), 
-                  dv.labels = "Perceived Listening Effort")
+                  dv.labels = "Perceived Listening Effort",
+                  file = "Tables/perceived_effort_model.html")
 
 ```
+
+```{r, echo = F, warning = F, message = F}
+
+rm(list = ls())
+
+```

### 2024-03-18 S. Bikulcius Thesis/analysis_report.Rmd
added standardized models

@@
 author: "Micah E. Hirsch"
-date: "2024-03-14"
+date: "2024-03-18"
 output: 
@@
 
+##### Unstandaridized Model
+
 ```{r, warning = F, message = F}
@@
 
+##### Standaridized Model
+
+```{r, warning = F, message = F}
+
+m7_standardized <- glmmTMB(scale(peak_pupil) ~ scale(trial_c) + speaker + scale(list_sort_c)*speaker + (1|subject), data = accurate_df)
+sjPlot::tab_model(m7_standardized, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Working Memory", 
+                                      "Working Memory * Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation",
+                  file = "Tables/Peak_pupil_dilation_standardized_model.html")
+
+```
+
 ```{r, echo = F, warning = F, message = F}
 
-rm(m0, m1, m2, m2b, m3, m3b, m4, m5, m6, m7, m8, m9, m10, pupil_des)
+rm(m0, m1, m2, m2b, m3, m3b, m4, m5, m6, m7, m7_standardized, m8, m9, m10, pupil_des)
 
@@
 
+##### Standaridized Model
+
+```{r}
+
+m10_standardized <- glmmTMB(scale(effort_rating) ~ scale(trial_c) + speaker + scale(list_sort_c)*speaker + 
+                scale(card_sort_c)*speaker + (trial_c|subject) + (trial_c|code), data = accurate_df)
+sjPlot::tab_model(m10_standardized, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", 
+                                      "Working Memory", "Cognitive Flexibility", "Working Memory * Speaker [ALS]",
+                                      "Cognitive Flexibility * Speaker [ALS]"), 
+                  dv.labels = "Perceived Listening Effort",
+                  file = "Tables/perceived_effort_model_standardized.html")
+
+```
+
 ```{r, echo = F, warning = F, message = F}

### 2024-03-21 S. Bikulcius Thesis/analysis_report.Rmd
Updated figures for Sam

@@
 author: "Micah E. Hirsch"
-date: "2024-03-18"
+date: "2024-03-21"
 output: 
@@
                                                "Native Hawaiian or Other Pacific Islander", "Native American or Alaska Native",
-                                               "Biracial or Multiracial", "Prefer not to answer", "Race not listed")),
+                                               "Biracial or Multiracial", "Race not listed", "Prefer not to answer")),
                 native_lang = factor(native_lang, c("American English", "Not American English"))) |>
   dplyr::rename(subject = id)
+
 ```
@@
 ```{r, warning = F, message = F}
+
+# Color Palette for Figures
+my_pal <- c("#033F63", "#2CB5AE", "#28666E", "#4BAA9B", "#49E9B9", "#57A773")
 
@@
   labs(x = "Score", y = "", color = "Cognitive Measure", fill = "Cognitive Measure") +
+  scale_color_manual(values = c(my_pal[3],my_pal[4], my_pal[5], my_pal[6])) +
+  scale_fill_manual(values = c(my_pal[3],my_pal[4], my_pal[5], my_pal[6])) +
   theme_classic()
@@
 
-For our analysis, it is important to consider the correlations between working memory, inhibitory control, cognitive flexibility, and processing speed.
+#### Overall Correlations
+
+For our analysis, it is important to consider the correlations between working memory, inhibitory control, cognitive flexibility, processing speed, peak pupil dilation, and perceived listening effort ratings.
 The figure below shows a correlation matrix.
@@
 Values closer to 1 or -1 suggest a stronger positive or negative correlation between the two cognitive measures, respectively.
+The effort measures were averaged per listener across the speakers.
 
@@
 There are also weak-to-moderate correlations between inhibitory control and cognitive flexibility and processing speed and working memory.
-This means the following:
+There are also weak postive correlations between inhibitory control and cognitive flexibility with peak pupil dilation.
+Perceived listening effort has a weak negative relationship with working memory and inhibitory control.This means the following:
 
@@
 -   Listeners with higher processing speed scores also tend to have higher working memory scores
-
-```{r, warning = F, message = F}
-
-cog |>
-  dplyr::select(ends_with("_c")) |>
-  dplyr::rename('Working Memory' = list_sort_c,
-                'Inhibitory Control' = flanker_c,
-                'Cognitive Flexibility' = card_sort_c,
-                'Processing Speed' = pattern_c) |>
-  cor(use = "pairwise.complete.obs") |>
-  ggcorrplot(
-    type = "lower",
-    outline.col = "white",
-    lab = TRUE,
-    insig = "blank",
-    legend.title = "Correlation"
-  )
-
-```
-
-# Speaker Characteristics
-
-In this study, we use speech samples from two speakers: a control speaker with neurotypical speech and a speaker with dysarthria secondary to Amyotropic Lateral Sclerosis (ALS).
-The mean speaker intelligibility is reported below.
-This was calculated using the listener responses.
-Although the speaker with ALS has a mean intelligibility of approximately 82%, about listeners still repeated approximately half of their trials incorrectly.
-When listening to the control speaker, there was a higher proportion of accurate responses.
+-   Listeners with higher inhibitory control scores tend to have higher peak pupil dilation values
+-   Listeners with higher cognitive flexibility scores tend to have higher peak pupil dilation values
+-   Listeners with higher working memory scores tend to rate their perceived listening effort lower overall
+-   Listeners with higher inhibitory control scores tend to rate perceived listening effort lower overall
+
+All other correlations in the matrix are very weak.
 
@@
   dplyr::mutate(effort_rating = ifelse(effort_rating < 0, NA, effort_rating))
+
+pupil1 <- dplyr::left_join(pupil, cog, by = "subject")
+
+overall_corr <- pupil1 |>
+  dplyr::filter(rep_acc == "accurate") |>
+  dplyr::select(subject, trial, speaker, code, peak_pupil, effort_rating, ends_with("_c")) |>
+  dplyr::group_by(subject, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
+  dplyr::summarize(pupil = mean(peak_pupil),
+                   effort = mean(effort_rating)) |>
+  ungroup() |>
+  dplyr::select(!subject) |>
+  dplyr::rename('Working Memory' = list_sort_c,
+                'Inhibitory Control' = flanker_c,
+                'Cognitive Flexibility' = card_sort_c,
+                'Processing Speed' = pattern_c,
+                'Peak Pupil Dilation' = pupil,
+                'Perceived Listening Effort' = effort) |>
+  cor(use = "pairwise.complete.obs") |>
+  ggcorrplot(
+    type = "lower",
+    outline.col = "white",
+    lab = TRUE,
+    insig = "blank",
+    legend.title = "Correlation",
+    title = "Overall"
+  )
+
+overall_corr
+
+```
+
+#### Correlations for the Control Speaker
+
+We are also looking at these same relationships for each speaker. The correlation matrix for these variables for the control speaker is shown below.
+
+There are some similar correlation patterns between the cognitive measures and both effort measures for the Control speaker compared to the overall matrix above.
+For instance, there correlations between peak pupil dilation and both cognitive flexibility and inhibitory control were still weak and positive.
+There was also a weak and negative correlation between inhibitory control and perceived listening effort.
+However, there are also some noticable differences.
+To start, there is a weak positive correlation between cognitive flexibility and perceived listening effort ratings.
+It is also interesting to note there is a stronger relationship between both effort measures (peak pupil dilation and perceived effort) for the control speaker (r = 0.26).
+This means the following:
+
+-   Listeners who had higher cognitive flexibility scores tend to have greater peak pupil dilation values when listening to the control speaker
+-   Listeners who had higher inhibitory control scores tend to have greater peak pupil dilation values when listening to the control speaker
+-   Listeners who had higher working memory scores tend to give the control speaker lower perceived listening effort ratings
+-   Listeners who had higher inhibitory control scores tend to give the control speaker lower perceived listening effort ratings
+-   Listeners who had higher cognitive flexibility scores tend to give the control speaker higher perceived listening effort ratings
+-   The weak positive relationship between peak pupil dilation and perceived listening effort ratings means that listeners who had greater peak pupil dilation while listening to the control speaker also tend to give the control speaker higher perceived listening effort ratings.
+Note that this relationship is stronger compared to when looking at the overall relationship between these measures.
+
+```{r, warning = F, message = F}
+
+control_corr <- pupil1 |>
+  dplyr::filter(rep_acc == "accurate") |>
+  dplyr::select(subject, speaker, trial, speaker, code, peak_pupil, effort_rating, ends_with("_c")) |>
+  dplyr::group_by(subject, speaker, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
+  dplyr::summarize(pupil = mean(peak_pupil),
+                   effort = mean(effort_rating)) |>
+  ungroup() |>
+  dplyr::filter(speaker == "Control") |>
+  dplyr::select(!c(subject, speaker)) |>
+  dplyr::rename('Working Memory' = list_sort_c,
+                'Inhibitory Control' = flanker_c,
+                'Cognitive Flexibility' = card_sort_c,
+                'Processing Speed' = pattern_c,
+                'Peak Pupil Dilation' = pupil,
+                'Perceived Listening Effort' = effort) |>
+  cor(use = "pairwise.complete.obs") |>
+  ggcorrplot(
+    type = "lower",
+    outline.col = "white",
+    lab = TRUE,
+    insig = "blank",
+    legend.title = "Correlation",
+    title = "Control Speaker"
+  )
+
+control_corr
+
+```
+
+#### Correlations for the ALS Speaker
+
+We also ran a separate correlation matrix so we can examine the relationship between these cognitive measures and effort measures while the participants were listening to the ALS speaker.
+
+Overall, all the correlations between the cognitive measures are the same as the other two matricies above.
+However, there are some differences noted in the correlations between the cognitive measures and each of the effort measures.
+To start, the correlation between working memory and peak pupil dilation for the ALS speaker is weak and positive.
+However, compared to the control speaker, this relationship is stronger for the ALS speaker than the control speaker.
+The correlations between peak pupil dilation and both inhibitory control and cognitive flexibility also are weaker for the ALS speaker compared to the control speaker.
+The correlations between perceived listening effort ratings and both inhibitory control and working memory are much weaker for the ALS speaker compared to the control speaker.
+However, the relationship between perceived listening effort and cognitive flexibility for the ALS speaker is weak-to-moderate and negative.
+The relationship between the two effort measures is also very small for the ALS speaker (pretty much negligible).
+This relationship was stronger for the control speaker.
+Overall, this means:
+
+-   Listeners who had greater working memory scores tend to have greater peak pupil dilation values when listening to the ALS speaker
+-   Listeners who had greater cognitive flexibility scores tend to give the ALS speaker lower perceived listening effort ratings
+
+```{r, warning = F, message = F}
+
+ALS_corr <- pupil1 |>
+  dplyr::filter(rep_acc == "accurate") |>
+  dplyr::select(subject, speaker, trial, speaker, code, peak_pupil, effort_rating, ends_with("_c")) |>
+  dplyr::group_by(subject, speaker, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
+  dplyr::summarize(pupil = mean(peak_pupil),
+                   effort = mean(effort_rating)) |>
+  ungroup() |>
+  dplyr::filter(speaker == "ALS") |>
+  dplyr::select(!c(subject, speaker)) |>
+  dplyr::rename('Working Memory' = list_sort_c,
+                'Inhibitory Control' = flanker_c,
+                'Cognitive Flexibility' = card_sort_c,
+                'Processing Speed' = pattern_c,
+                'Peak Pupil Dilation' = pupil,
+                'Perceived Listening Effort' = effort) |>
+  cor(use = "pairwise.complete.obs") |>
+  ggcorrplot(
+    type = "lower",
+    outline.col = "white",
+    lab = TRUE,
+    insig = "blank",
+    legend.title = "Correlation",
+    title = "ALS Speaker"
+  )
+
+ALS_corr
+
+```
+
+```{r, eval = F, message = F, warning = F}
+
+corr_plot <- ggpubr::ggarrange(overall_corr, control_corr, ALS_corr, ncol = 3, legend = "bottom", common.legend = T)
+
+ggsave("Figures/Correlation_matricies.png", plot = corr_plot, width = 6, height = 3, units = "in", scale = 2)
+
+rm(ALS_corr, control_corr, overall_corr)
+
+```
+
+# Speaker Characteristics
+
+In this study, we use speech samples from two speakers: a control speaker with neurotypical speech and a speaker with dysarthria secondary to Amyotropic Lateral Sclerosis (ALS).
+The mean speaker intelligibility is reported below.
+This was calculated using the listener responses.
+Although the speaker with ALS has a mean intelligibility of approximately 82%, about listeners still repeated approximately half of their trials incorrectly.
+When listening to the control speaker, there was a higher proportion of accurate responses.
+
+```{r, warning = F, message = F}
 
@@
   theme_classic() +
+  scale_color_manual(values = my_pal) +
+  scale_fill_manual(values = my_pal) +
   theme(aspect.ratio = 1)
+
+ggsave("Figures/peak_pupil_accurate_plot.png", plot = last_plot())
 
@@
   theme_classic() +
+  scale_color_manual(values = my_pal) +
+  scale_fill_manual(values = my_pal) +
   theme(aspect.ratio = 1)
+
+ggsave("Figures/peak_pupil_accurate_inaccurate_plot.png", plot = last_plot())
 
@@
 ```{r, warning = F, message = F}
-
-pupil1 <- dplyr::left_join(pupil, cog, by = "subject")
 
@@
       color = speaker,
-      ) +
-  geom_point(alpha = 0.3) +
+      shape = speaker) +
+  geom_point(alpha = .6) +
   geom_smooth(method = "lm") +
@@
   theme_classic() +
+  scale_color_manual(values = my_pal) +
+  scale_shape_manual(values = c(1, 2)) +
   theme(aspect.ratio = 1)
+
+ggsave("Figures/cog_peak_pupil_accurate_plot.png", plot = last_plot())
 
@@
   theme_classic() +
+  scale_color_manual(values = my_pal) +
+  scale_shape_manual(values = c(1, 2)) +
   theme(aspect.ratio = 1)
 
-ggsave("Figures/peak_pupil_cog_plot.png", plot = last_plot())
+ggsave("Figures/cog_peak_pupil_acc_inacc_plot.png", plot = last_plot())
 
@@
 
-##### Unstandaridized Model
-
 ```{r, warning = F, message = F}
@@
 
-##### Standaridized Model
-
-```{r, warning = F, message = F}
+```{r, eval = F, warning = F, message = F}
 
@@
   theme_classic() +
+  scale_color_manual(values = my_pal) +
+  scale_fill_manual(values = my_pal) +
   theme(aspect.ratio = 1)
+
+ggsave("Figures/effort_rating_accurate_plot.png", plot = last_plot())
 
@@
   theme_classic() +
+  scale_color_manual(values = my_pal) +
+  scale_fill_manual(values = my_pal) +
   theme(aspect.ratio = 1)
+
+ggsave("Figures/effort_rating_acc_inacc_plot.png", plot = last_plot())
 
@@
       color = speaker,
-      ) +
-  geom_point(alpha = 0.3) +
+      shape = speaker) +
+  geom_point(alpha = 0.6) +
   geom_smooth(method = "lm") +
@@
   theme_classic() +
+  scale_color_manual(values = my_pal) +
+  scale_shape_manual(values = c(1, 2)) +
   theme(aspect.ratio = 1)
+
+ggsave("Figures/cog_effort_rating_accurate_plot.png", plot = last_plot())
 
@@
   theme_classic() +
+  scale_color_manual(values = my_pal) +
+  scale_shape_manual(values = c(1, 2)) +
   theme(aspect.ratio = 1)
+
+ggsave("Figures/cog_effort_rating_acc_inacc_plot.png", plot = last_plot())
 
@@
 
-##### Standaridized Model
-
-```{r}
+```{r, eval = F, warning = F, message = F}
 

### 2024-07-24 S. Bikulcius Thesis/analysis_report.Rmd

@@
   scale_fill_manual(values = c(my_pal[3],my_pal[4], my_pal[5], my_pal[6])) +
-  theme_classic()
+  theme_classic() +
+  theme(legend.position = "none")
+
+ggsave("Figures/cog_distribution.png", plot = last_plot())
 

### 2024-07-24 Manuscript Analysis/manuscript_analysis.Rmd
cleaned cognitive data

---
title: "Cognitive Analysis"
author: "Micah E. Hirsch"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

The purpose of this document is to share the analysis code and report the results for the manuscript. The full Rmd version of the document can be found on the project's Github repository or OSF project page. The data cleaning code for the pupil and perceived listening effort ratings can be found in the Github repo. Analyses were conducted using R version 4.3.3.

```{r}

# Loading in the needed packages

library(tidyverse)
library(rio)
library(glmmTMB)
library(janitor)
library(ggpubr)
library(patchwork)
library(gt)
library(gtsummary)
library(ggridges)
library(performance)
library(sjPlot)
library(ggcorrplot)

```

#

```{r}

plot(pressure)

```

### 2024-07-24 Manuscript Analysis/data_prep.R

@@ -1,0 +1,53 @@
+# Data Processing
+
+# Author: Micah E. Hirsch, Ph.D. (mhirsch@fsu.edu)
+
+# Data: 7/24/2024
+
+# Purpose: To load and format cognitive data and merge it with the cleaned 
+# data from https://github.com/micah-hirsch/Listening-Effort-in-Dysarthria
+
+# Loading Needed Packages
+
+library(tidyverse)
+library(rio)
+library(janitor)
+
+# Setting Working Directory
+
+setwd("~/Documents/LE-Cognitive-Study/Raw Data")
+
+file <- list.files()
+
+file <- file[grepl("Assessment Scores", file)]
+
+cog_data <- rio::import(file) |>
+  janitor::clean_names() |>
+  dplyr::select(pin, inst, uncorrected_standard_score, age_corrected_standard_score) |>
+  dplyr::mutate(measure = case_when(grepl("working memory", inst, ignore.case = T) ~ "list_sort",
+                                    grepl("flanker", inst, ignore.case = T) ~ "flanker",
+                                    grepl("dimensional change", inst, ignore.case = T) ~ "card_sort",
+                                    TRUE ~ "pattern"), .after = inst) |>
+  dplyr::select(-inst) |>
+  tidyr::pivot_longer(cols = uncorrected_standard_score:age_corrected_standard_score,
+                      names_to = "type",
+                      values_to = "score") |>
+  dplyr::mutate(measure = ifelse(grepl("uncorrected", type, ignore.case = T), 
+                                  paste(measure, "u", sep = "_"), paste(measure, "c", sep = "_"))) |>
+  dplyr::select(-c(type)) |>
+  tidyr::pivot_wider(names_from = measure,
+                     values_from = score) |>
+  dplyr::rename(id = pin)
+
+rm(file)
+
+# Importing demographic data
+
+setwd("~/Documents/LE-Cognitive-Study/Manuscript Analysis/Cleaned Data")
+
+demo <- rio::import("cleaned_listener_demo.csv")
+
+demo <- demo |>
+  dplyr::left_join(cog_data, by = "id")
+
+rio::export(cog_data, "cleaned_cog_data.csv")

### 2024-07-25 Manuscript Analysis/manuscript_analysis.Rmd
Listener demo and cog summary

@@
 
+# Load in the listener demographics data
+
+demo <- rio::import("Cleaned Data/cleaned_listener_demo.csv") |>
+  dplyr::filter(pupil_complete == "complete") |>
+  dplyr::select(id, age, gender, ethnicity, race, native_lang) |>
+  dplyr::mutate(gender = factor(gender, levels = c("Man", "Woman", "Nonbinary", "Questioning", "Prefer not to answer")),
+                ethnicity = factor(ethnicity, levels = c("Hispanic/Latino(a/e)", "Not Hispanic/Latino(a/e)", "Prefer not to answer")),
+                race = factor(race, levels = c("white/Caucasian", "Black/African American", "Asian/Asian American",
+                                               "Native Hawaiian or Other Pacific Islander", "Native American or Alaska Native",
+                                               "Biracial or Multiracial", "Race not listed", "Prefer not to answer")),
+                native_lang = factor(native_lang, c("American English", "Not American English"))) |>
+  dplyr::rename(subject = id)
+
+# Loading in the cog data
+
+cog <- rio::import("Cleaned Data/cleaned_cog_data.csv")
+
+# Loading in the pupil data
+
+pupil <- rio::import("Cleaned Data/cleaned_pupil_data_normalized.csv")
+
+# Loading in the phrase accuracy data
+
+phrase_acc <- rio::import("Cleaned Data/repetition_accuracy.csv") |>
+  dplyr::mutate(rep_acc = ifelse(initial_response == "missing data", NA, 
+                                 ifelse(target_number == correct_words_initial, "accurate", "inaccurate")),
+                rep_acc_rel = ifelse(rel_response == "missing data", NA, 
+                                 ifelse(target_number == correct_words_rel, "accurate", "inaccurate")))
+
 ```
 
-#
+# Listener Demographics
+
+The listeners used in this study are the same from the parent study. In brief, the demographic information from these listeners are below.
 
@@
 
-plot(pressure)
+demo_table <- demo |>
+  dplyr::select(age, gender, ethnicity, race, native_lang) |>
+  tbl_summary(type = list(age ~ "continuous",
+                          gender ~ "categorical",
+                          ethnicity ~ "categorical",
+                          race ~ "categorical",
+                          native_lang ~ "categorical"),
+              statistic = list(all_continuous() ~ "{mean} ({sd})",
+                               all_categorical() ~ "{n} ({p}%)"),
+              digits = list(everything() ~ c(2)),
+              label = list(age ~ "Age",
+                           gender ~ "Gender",
+                           ethnicity ~ "Ethnicity",
+                           race ~ "Race",
+                           native_lang ~ "Native Language")) |>
+  as_gt()
+
+demo_table
+
+demo_table |>
+  gt::gtsave("Tables/demo_table.html")
 
 ```
+
+# Listener Cognition
+
+```{r}
+
+cog_table <- cog |>
+  dplyr::select(-subject) |>
+  # Just using the age-corrected scores
+  dplyr::select(ends_with("_c")) |>
+  tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})"),
+              digits = list(everything() ~ c(2)),
+              label = list(list_sort_c ~ "Working Memory",
+                           flanker_c ~ "Inhibitory Control",
+                           card_sort_c ~ "Cognitive Flexibility",
+                           pattern_c ~ "Processing Speed")) |>
+  as_gt()
+
+cog_table
+
+cog_table |>
+  gt::gtsave("Tables/cog_table.html")
+
+```

### 2024-07-25 Manuscript Analysis/data_prep.R

@@ -37,17 +37,12 @@ cog_data <- rio::import(file) |>
   dplyr::select(-c(type)) |>
   tidyr::pivot_wider(names_from = measure,
                      values_from = score) |>
-  dplyr::rename(id = pin)
+  dplyr::rename(subject = pin)
 
 rm(file)
 
 # Importing demographic data
 
 setwd("~/Documents/LE-Cognitive-Study/Manuscript Analysis/Cleaned Data")
-
-demo <- rio::import("cleaned_listener_demo.csv")
-
-demo <- demo |>
-  dplyr::left_join(cog_data, by = "id")
 
 rio::export(cog_data, "cleaned_cog_data.csv")

### 2024-07-25 Manuscript Analysis/manuscript_analysis.Rmd
WM pupil analysis

@@
 library(ggcorrplot)
+library(mgcv) # install.packages("mgcv")
+library(itsadug) # install.packages("itsadug")
 
@@
 
-# Listener Demographics
+# Listener Characteristics
 
-The listeners used in this study are the same from the parent study. In brief, the demographic information from these listeners are below.
+## Demographic Information
+
+The listeners used in this study are the same from the parent study.
+In brief, the demographic information from these listeners are below.
+A total of 34 young adult listeners participated in the experiment.
 
@@
 
-# Listener Cognition
+## Listener Cognition
+
+The information below reports the mean age-corrected standardized scores on the subtests from the NIH Toolbox cognition battery (working memory, inhibitory control, cognitive flexibility, and processing speed).
 
@@
 ```
+
+```{r}
+
+# Color Palette for Figures
+my_pal <- c("#033F63", "#2CB5AE", "#28666E", "#4BAA9B", "#49E9B9", "#57A773")
+
+cog |>
+  dplyr::select(subject, ends_with("_c")) |>
+  dplyr::rename("Working Memory" = list_sort_c,
+                "Inhibitory Control" = flanker_c,
+                "Cognitive Flexibility" = card_sort_c,
+                "Processing Speed" = pattern_c) |>
+  tidyr::pivot_longer(cols = 'Working Memory':'Processing Speed',
+                      names_to = "cog_measure",
+                      values_to = "score") |>
+  ggplot() +
+  aes(x = score,
+      y = cog_measure,
+      color = cog_measure,
+      fill = cog_measure) +
+  geom_density_ridges(jittered_points = T,
+                      position = position_points_jitter(width = 0.05, height = 0),
+                      point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.5) +
+  labs(x = "Score", y = "", color = "Cognitive Measure", fill = "Cognitive Measure") +
+  scale_color_manual(values = c(my_pal[3],my_pal[4], my_pal[5], my_pal[6])) +
+  scale_fill_manual(values = c(my_pal[3],my_pal[4], my_pal[5], my_pal[6])) +
+  theme_classic() +
+  theme(legend.position = "none")
+
+```
+
+```{r}
+
+# Removing unneeded items
+
+rm(cog_table, demo_table, demo)
+
+```
+
+# Pupil Dilation
+
+```{r}
+
+# Selecting needed variables from phrase repetition accuracy df
+
+acc <- phrase_acc |>
+  dplyr::select(subject, trial, target_number, correct_words_initial, rep_acc)
+
+## Merging acc df with pupil dilation df
+
+pupil <- pupil |>
+  # Merging phrase accuracy information with df
+  dplyr::left_join(acc, by = c("subject", "trial")) |>
+  # Filter out trials with missing phrase repetition data |>
+  dplyr::filter(!is.na(rep_acc)) |>
+  # Merging cognitive data
+  dplyr::left_join(cog, by = "subject")
+
+```
+
+## Base Model
+
+Final Model from parent study
+
+```{r}
+
+pupil <- pupil |>
+  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
+                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")),
+                phrase = as.factor(code),
+                listener = as.factor(subject),
+                condition = case_when(speaker == "Control" & rep_acc == "accurate" ~ "Control Accurate",
+                                      speaker == "Control" & rep_acc == "inaccurate" ~ "Control Inaccurate",
+                                      speaker == "ALS" & rep_acc == "accurate" ~ "ALS Accurate",
+                                      TRUE ~ "ALS Inaccurate"),
+                condition = factor(condition, 
+                                   levels = c("Control Accurate", "Control Inaccurate", "ALS Accurate", "ALS Inaccurate"))) |>
+  dplyr::group_by(listener, speaker) |>
+  dplyr::mutate(num_points = n()) |>
+  dplyr::mutate(start_event = c(TRUE, rep(FALSE, each = (num_points - 1)))) |>
+  ungroup()
+
+m0_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  # Random Smooth for listener
+                  s(time_norm, listener, bs = 'fs', m = 1),
+                discrete = T,
+                AR.start = start_event,
+                rho = .9341941,
+                family = "scat",
+                data = pupil)
+
+summary.gam(m0_pupil)
+
+m0_prediction <- plot_smooth(m0_pupil,
+            view = "time_norm",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T)
+
+```
+
+```{r}
+
+m0_predicted_values <- m0_prediction[[1]][c("condition", "time_norm", "fit", "ul", "ll")] |>
+  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
+                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"),
+                speaker = factor(speaker, levels = c("Control", "ALS")))
+
+m0_predicted_values |>
+  ggplot() +
+  aes(x = time_norm,
+      y = fit,
+      color = speaker,
+      fill = speaker) +
+  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
+  geom_line() +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.15) +
+  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
+  facet_wrap("rep_acc") +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom") +
+  labs(x = "Normalized Time (ms)",
+       y = "Pupil Dilation (Arbitrary Units)",
+       color = "Speaker",
+       fill = "Speaker",
+       linetype = "Speaker") +
+  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS"))
+
+```
+
+## Working Memory
+
+### Initial Model
+
+```{r}
+
+wm_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  te(time_norm, list_sort_c, k = 20) +
+                  # Random Smooth for listener
+                  s(time_norm, listener, bs = 'fs', m = 1),
+                discrete = T,
+                AR.start = start_event,
+                rho = .9341941,
+                family = "scat",
+                data = pupil)
+
+summary.gam(wm_pupil)
+
+```
+
+#### Model Check
+
+```{r}
+
+gam.check(wm_pupil)
+
+acf_resid(wm_pupil)
+
+```
+
+#### Model Comparison
+
+```{r}
+
+compareML(m0_pupil, wm_pupil, signif.stars = T, suggest.report = T)
+
+```
+
+### Model 1
+
+```{r}
+
+wm_pupil_1 <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  te(time_norm, list_sort_c, by = speaker, k = 20) +
+                  # Random Smooth for listener
+                  s(time_norm, listener, bs = 'fs', m = 1),
+                discrete = T,
+                AR.start = start_event,
+                rho = .9341941,
+                family = "scat",
+                data = pupil)
+
+summary.gam(wm_pupil_1)
+
+```
+
+#### Model Check
+
+```{r}
+
+gam.check(wm_pupil_1)
+
+acf_resid(wm_pupil_1)
+
+```
+
+#### Model Comparison
+
+```{r}
+
+compareML(m0_pupil, wm_pupil_1, signif.stars = T, suggest.report = T)
+
+```
+### Model 2
+
+```{r}
+
+wm_pupil_2 <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  te(time_norm, list_sort_c, by = condition, k = 20) +
+                  # Random Smooth for listener
+                  s(time_norm, listener, bs = 'fs', m = 1),
+                discrete = T,
+                AR.start = start_event,
+                rho = .9341941,
+                family = "scat",
+                data = pupil)
+
+summary.gam(wm_pupil_2)
+
+```
+
+#### Model Check
+
+```{r}
+
+gam.check(wm_pupil_2)
+
+acf_resid(wm_pupil_2)
+
+```
+#### Model Comparison
+
+```{r}
+
+compareML(m0_pupil, wm_pupil_2, signif.stars = T, suggest.report = T)
+
+```

### 2024-07-25 Manuscript Analysis/manuscript_analysis.Rmd
wm model

@@
 
-### Initial Model
-
 ```{r}
@@
                   s(time_norm, by = condition, k = 25) +
-                  te(time_norm, list_sort_c, k = 20) +
+                  s(list_sort_c, by = condition, k = 20) +
                   # Random Smooth for listener
@@
 summary.gam(wm_pupil)
+
+plot_diff(wm_pupil, view = "list_sort_c",
+          comp = list(condition = c("ALS Inaccurate", "ALS Accurate")))
+
+plot_diff(wm_pupil, view = "list_sort_c",
+          comp = list(condition = c("Control Inaccurate", "Control Accurate")))
+
+plot_diff(wm_pupil, view = "list_sort_c",
+          comp = list(condition = c("ALS Accurate", "Control Accurate")))
 
@@
 
-### Model 1
-
 ```{r}
 
-wm_pupil_1 <- bam(normed_pupil ~ condition +
-                  s(time_norm, by = condition, k = 25) +
-                  te(time_norm, list_sort_c, by = speaker, k = 20) +
-                  # Random Smooth for listener
-                  s(time_norm, listener, bs = 'fs', m = 1),
-                discrete = T,
-                AR.start = start_event,
-                rho = .9341941,
-                family = "scat",
-                data = pupil)
+wm_pred <- expand_grid(
+  time_norm = seq(from=min(pupil$time_norm), 
+              to=max(pupil$time_norm), 
+              length.out = 100),
+  list_sort_c = seq(from=min(pupil$list_sort_c), 
+              to=max(pupil$list_sort_c), 
+              length.out = 100),
+  condition = c("Control Accurate", "Control Inaccurate", "ALS Accurate", "ALS Inaccurate"),
+  listener = c("LE01", "LE02", "LE04", "LE05", "LE06", "LE07", "LE08", "LE09", "LE10", "LE11",
+               "LE12", "LE13", "LE15", "LE16", "LE18", "LE19", "LE20", "LE21", "LE22", "LE23",
+               "LE24", "LE25", "LE26", "LE27", "LE28", "LE29", "LE30", "LE31", "LE32", "LE33", 
+               "LE34", "LE35", "LE36", "LE38")
+)
 
-summary.gam(wm_pupil_1)
+wm_pred <- predict(wm_pupil, newdata = wm_pred, 
+                      se.fit = TRUE) |> 
+  as_tibble() |>
+  cbind(wm_pred)
 
 ```
-
-#### Model Check
-
-```{r}
-
-gam.check(wm_pupil_1)
-
-acf_resid(wm_pupil_1)
-
-```
-
-#### Model Comparison
-
-```{r}
-
-compareML(m0_pupil, wm_pupil_1, signif.stars = T, suggest.report = T)
-
-```
-### Model 2
-
-```{r}
-
-wm_pupil_2 <- bam(normed_pupil ~ condition +
-                  s(time_norm, by = condition, k = 25) +
-                  te(time_norm, list_sort_c, by = condition, k = 20) +
-                  # Random Smooth for listener
-                  s(time_norm, listener, bs = 'fs', m = 1),
-                discrete = T,
-                AR.start = start_event,
-                rho = .9341941,
-                family = "scat",
-                data = pupil)
-
-summary.gam(wm_pupil_2)
-
-```
-
-#### Model Check
-
-```{r}
-
-gam.check(wm_pupil_2)
-
-acf_resid(wm_pupil_2)
-
-```
-#### Model Comparison
-
-```{r}
-
-compareML(m0_pupil, wm_pupil_2, signif.stars = T, suggest.report = T)
-
-```

### 2024-07-26 Manuscript Analysis/manuscript_analysis.Rmd
initial analyses

--- 2024-07-25 Manuscript Analysis/manuscript_analysis.Rmd
@@
 library(itsadug) # install.packages("itsadug")
+library(viridis) # install.packages("viridis")
 
@@
 pupil <- rio::import("Cleaned Data/cleaned_pupil_data_normalized.csv")
+
+# Loading in the PLE data
+
+ple <- rio::import("Cleaned Data/cleaned_ple_data.csv")
 
@@
 
-Final Model from parent study
+Final Model from parent study. This is just modeling pupil dilation changing by levels of condition (interaction between speaker and phrase repetition accuracy).
 
@@
 
-### Initial Model
+Now, to test whether pupil dilation is impacted by working memory capacity, a smooth was added to this model: the age-corrected standardized working memory scores from the NIH Toolbox Cognition battery.
+The smooth was added per level of "condition" (interaction between speaker and repetition accuracy).
 
@@
                   s(time_norm, by = condition, k = 25) +
-                  te(time_norm, list_sort_c, k = 20) +
+                  s(list_sort_c, by = condition) +
                   # Random Smooth for listener
@@
 
+wm_prediction <- plot_smooth(wm_pupil,
+            view = "list_sort_c",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T)
+
 ```
@@
 #### Model Check
+
+Checking model assumptions by plotting residuals and autocorrelation plots. Overall, the plots indicate the model is a good fit to the data.
 
@@
 ```
-
-### Model 1
+### Working Memory vs Pupil Dilation
 
@@
 
-wm_pupil_1 <- bam(normed_pupil ~ condition +
+wm_predicted_values <- wm_prediction[[1]][c("condition", "list_sort_c", "fit", "ul", "ll")] |>
+  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
+                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"),
+                speaker = factor(speaker, levels = c("Control", "ALS")))
+
+wm_predicted_values |>
+  ggplot() +
+  aes(x = list_sort_c,
+      y = fit,
+      color = rep_acc,
+      fill = rep_acc) +
+  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
+  geom_line() +
+  facet_wrap("speaker") +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom") +
+  labs(x = "Working Memory",
+       y = "Pupil Dilation (Arbitrary Units)",
+       color = "Repetition Accuracy",
+       fill = "Repetition Accuracy",
+       linetype = "Repetition Accuracy") +
+  scale_color_manual(values = c(my_pal[1], my_pal[2])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[2]))
+
+```
+
+## Inhibitory Control
+
+```{r}
+
+inhib_pupil <- bam(normed_pupil ~ condition +
                   s(time_norm, by = condition, k = 25) +
-                  te(time_norm, list_sort_c, by = speaker, k = 20) +
+                  s(flanker_c, by = condition) +
                   # Random Smooth for listener
@@
 
-summary.gam(wm_pupil_1)
+summary.gam(inhib_pupil)
+
+inhib_prediction <- plot_smooth(inhib_pupil,
+            view = "flanker_c",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T)
 
@@
 
-gam.check(wm_pupil_1)
+gam.check(inhib_pupil)
 
-acf_resid(wm_pupil_1)
+acf_resid(inhib_pupil)
 
@@
 
-compareML(m0_pupil, wm_pupil_1, signif.stars = T, suggest.report = T)
+compareML(m0_pupil, inhib_pupil, signif.stars = T, suggest.report = T)
 
 ```
-### Model 2
+
+### Inhibitory Control vs Pupil Dilation
 
@@
 
-wm_pupil_2 <- bam(normed_pupil ~ condition +
+inhib_predicted_values <- inhib_prediction[[1]][c("condition", "flanker_c", "fit", "ul", "ll")] |>
+  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
+                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"),
+                speaker = factor(speaker, levels = c("Control", "ALS")))
+
+inhib_predicted_values |>
+  ggplot() +
+  aes(x = flanker_c,
+      y = fit,
+      color = rep_acc,
+      fill = rep_acc) +
+  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
+  geom_line() +
+  facet_wrap("speaker") +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom") +
+  labs(x = "Inhibitory Control",
+       y = "Pupil Dilation (Arbitrary Units)",
+       color = "Repetition Accuracy",
+       fill = "Repetition Accuracy",
+       linetype = "Repetition Accuracy") +
+  scale_color_manual(values = c(my_pal[1], my_pal[2])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[2]))
+
+```
+
+## Cognitive Flexibility
+
+```{r}
+
+flex_pupil <- bam(normed_pupil ~ condition +
                   s(time_norm, by = condition, k = 25) +
-                  te(time_norm, list_sort_c, by = condition, k = 20) +
+                  s(card_sort_c, by = condition, k = 20) +
                   # Random Smooth for listener
@@
 
-summary.gam(wm_pupil_2)
+summary.gam(flex_pupil)
+
+flex_prediction <- plot_smooth(flex_pupil,
+            view = "card_sort_c",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T)
 
@@
 
-#### Model Check
+### Model Check
 
@@
 
-gam.check(wm_pupil_2)
+gam.check(flex_pupil)
 
-acf_resid(wm_pupil_2)
+acf_resid(flex_pupil)
 
 ```
-#### Model Comparison
+
+### Model Comparison
 
@@
 
-compareML(m0_pupil, wm_pupil_2, signif.stars = T, suggest.report = T)
+compareML(m0_pupil, flex_pupil, signif.stars = T, suggest.report = T)
 
 ```
+
+### Cognitive Flexibility vs Pupil Dilation
+
+```{r}
+
+flex_predicted_values <- flex_prediction[[1]][c("condition", "card_sort_c", "fit", "ul", "ll")] |>
+  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
+                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"),
+                speaker = factor(speaker, levels = c("Control", "ALS")))
+
+flex_predicted_values |>
+  ggplot() +
+  aes(x = card_sort_c,
+      y = fit,
+      color = rep_acc,
+      fill = rep_acc) +
+  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
+  geom_line() +
+  facet_wrap("speaker") +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom") +
+  labs(x = "Cognitive Flexibility",
+       y = "Pupil Dilation (Arbitrary Units)",
+       color = "Repetition Accuracy",
+       fill = "Repetition Accuracy",
+       linetype = "Repetition Accuracy") +
+  scale_color_manual(values = c(my_pal[1], my_pal[2])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[2]))
+
+```
+
+## Processing Speed
+
+```{r}
+
+process_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  s(pattern_c, by = condition, k = 15) +
+                  # Random Smooth for listener
+                  s(time_norm, listener, bs = 'fs', m = 1),
+                discrete = T,
+                AR.start = start_event,
+                rho = .9341941,
+                family = "scat",
+                data = pupil)
+
+summary.gam(process_pupil)
+
+process_prediction <- plot_smooth(process_pupil,
+            view = "pattern_c",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T)
+
+```
+
+### Model Check
+
+```{r}
+
+gam.check(process_pupil)
+
+acf_resid(process_pupil)
+
+```
+
+### Model Comparison
+
+```{r}
+
+compareML(m0_pupil, process_pupil, signif.stars = T, suggest.report = T)
+
+```
+
+### Processing Speed vs Pupil Dilation
+
+```{r}
+
+process_predicted_values <- process_prediction[[1]][c("condition", "pattern_c", "fit", "ul", "ll")] |>
+  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
+                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"),
+                speaker = factor(speaker, levels = c("Control", "ALS")))
+
+process_predicted_values |>
+  ggplot() +
+  aes(x = pattern_c,
+      y = fit,
+      color = rep_acc,
+      fill = rep_acc) +
+  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
+  geom_line() +
+  facet_wrap("speaker") +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom") +
+  labs(x = "Processing Speed",
+       y = "Pupil Dilation (Arbitrary Units)",
+       color = "Repetition Accuracy",
+       fill = "Repetition Accuracy",
+       linetype = "Repetition Accuracy") +
+  scale_color_manual(values = c(my_pal[1], my_pal[2])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[2]))
+
+```
+
+# Perceived Listening Effort Ratings
+
+## Visualization
+
+```{r}
+
+# Merge PLE data with rep acc and listener cog
+
+ple <- ple |>
+  dplyr::left_join(acc, by = c("subject", "trial")) |>
+  dplyr::filter(!is.na(rep_acc)) |>
+  dplyr::left_join(cog, by = "subject") |>
+  dplyr::filter(effort_rating > 0) |>
+  dplyr::filter(effort_rating <= 7) |>
+  dplyr::mutate(trial_c = trial - 6,
+                speaker = factor(speaker, levels = c("Control", "ALS")))
+
+ple |>
+  dplyr::select(subject, trial, speaker, code, effort_rating, rep_acc, ends_with("_c")) |>
+  dplyr::group_by(subject, speaker, rep_acc, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
+  dplyr::summarize(effort = mean(effort_rating)) |>
+  dplyr::rename('Working Memory' = list_sort_c,
+                'Inhibitory Control' = flanker_c,
+                'Cognitive Flexibility' = card_sort_c,
+                'Processing Speed' = pattern_c) |>
+  tidyr::pivot_longer(cols = "Working Memory":"Processing Speed",
+                     names_to = "cog_measure",
+                     values_to = "score") |>
+  ungroup() |>
+  ggplot() +
+  aes(x = score,
+      y = effort,
+      color = speaker,
+      shape = speaker) +
+  geom_point(alpha = 0.6) +
+  geom_smooth(method = "lm") +
+  facet_grid(rep_acc ~ cog_measure) +
+  labs(x = "Score", y = "Perceived Listening Effort Rating") +
+  theme_classic() +
+  scale_color_manual(values = my_pal) +
+  scale_shape_manual(values = c(1, 2)) +
+  theme(aspect.ratio = 1)
+
+```
+
+## Models
+
+### Fully Unconditional Model
+
+```{r, warning = F, message = F}
+
+m0 <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = ple)
+sjPlot::tab_model(m0, pred.labels = "Intercept",
+                  dv.labels = "Perceived Listening Effort")
+
+performance::icc(m0)
+
+```
+
+### Model 1
+
+In this step, we are adding trial order (trial_c) as a random slop for both the listener (subject) and phrase (code) random intercepts.
+The random intercept of trial order significantly improved model fit.
+
+```{r, warning = F, message = F}
+
+m1<- glmmTMB(effort_rating ~ 1 + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m1, pred.labels = "Intercept",
+                  dv.labels = "Perceived Listening Effort")
+performance::test_performance(m0, m1)
+
+```
+
+### Model 2
+
+Trial order is the first random effect added to the model.
+Trial order is added in order to control for any possible order effects in the data.
+As noted in the results below, adding trial order did not significantly improve model fit to the data.
+However, I'll leave this fixed effect in the model to act as a covariate.
+
+```{r, warning = F, message = F}
+
+m2 <- glmmTMB(effort_rating ~ trial_c + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m2, 
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m1, m2)
+
+```
+
+### Model 3
+
+Adding the main fixed effect of speaker significantly improved the model fit.
+
+```{r, warning = F, message = F}
+
+m3 <- glmmTMB(effort_rating ~ trial_c + speaker + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m3,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m2, m3)
+
+```
+
+### Model 4
+
+The main fixed effect of repetition accuracy significantly improved model fit.
+
+```{r, warning = F, message = F}
+
+m4 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m4,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m3, m4)
+
+```
+### Model 5
+
+The interaction term between speaker and repetition accuracy did not significantly improve model fit.
+This indicates that the magnitude between accurate and inaccurate ratings was similar for both speakers.
+
+```{r, warning = F, message = F}
+
+m5 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m5,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m4, m5)
+
+```
+### Model 6
+
+```{r}
+
+m6 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + scale(list_sort_c, scale = F) + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m6,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m4, m6)
+
+```
+### Model 7
+
+```{r}
+
+m7 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + scale(flanker_c, scale = F) + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m7,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m4, m7)
+
+```
+### Model 8
+
+```{r}
+
+m8 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + scale(card_sort_c, scale = F) + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m8,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m4, m8)
+
+```
+
+### Model 9
+
+```{r}
+
+m9 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + scale(pattern_c, scale = F) + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m9,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m4, m9)
+
+```
+### Model 10
+
+```{r}
+
+m10 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*scale(list_sort_c, scale = F) + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m10,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m4, m10)
+
+```
+
+### Model 11
+
+```{r}
+
+m11 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*scale(flanker_c, scale = F) + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m11,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m4, m11)
+
+```
+
+### Model 12
+
+```{r}
+
+m12 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*scale(card_sort_c, scale = F) + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m12,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m4, m12)
+
+```
+### Model 13
+
+```{r}
+
+m13 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                 speaker*scale(card_sort_c, scale = F) +  speaker*scale(pattern_c, scale = F) +(trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m13,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m12, m13)
+
+```
+
+### Model 14
+
+```{r}
+
+m14 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                 speaker*scale(card_sort_c, scale = F) +  speaker*scale(list_sort_c, scale = F) + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m14,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m12, m14)
+
+```
+
+### Model 15
+
+```{r}
+
+m15 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                 speaker*scale(card_sort_c, scale = F) +  speaker*scale(list_sort_c, scale = F) + 
+                 rep_acc*scale(list_sort_c, scale = F) +
+                 (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m15,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m14, m15)
+
+```
+### Model 16
+
+```{r}
+
+m16 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                 speaker*scale(card_sort_c, scale = F) + speaker*scale(list_sort_c, scale = F) + 
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(flanker_c, scale = F) +
+                 (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m16,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m15, m16)
+
+```
+### Model 17
+
+```{r}
+
+m17 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                 speaker*scale(card_sort_c, scale = F) + speaker*scale(list_sort_c, scale = F) + 
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
+                 (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m17,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m15, m17)
+
+```
+
+### Model 18
+
+```{r}
+
+m18 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                 speaker*scale(card_sort_c, scale = F) + speaker*scale(list_sort_c, scale = F) + 
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) + rep_acc*scale(pattern_c, scale = F) +
+                 (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m18,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m17, m18)
+
+```
+
+### Model 19
+
+```{r}
+
+m19 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                 speaker*scale(card_sort_c, scale = F) + speaker*scale(list_sort_c, scale = F) + 
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) + 
+                 speaker*rep_acc*scale(list_sort_c, scale = F) +
+                 (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m19,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m17, m19)
+
+```
+### Model 20
+
+```{r}
+
+m20 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                 speaker*scale(card_sort_c, scale = F) + speaker*scale(list_sort_c, scale = F) + 
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) + 
+                 speaker*rep_acc*scale(flanker_c, scale = F) +
+                 (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m20,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m17, m20)
+
+```
+### Model 21
+
+```{r}
+
+m21 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                 speaker*scale(card_sort_c, scale = F) + speaker*scale(list_sort_c, scale = F) + 
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) + 
+                 speaker*rep_acc*scale(card_sort_c, scale = F) +
+                 (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m21,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m17, m21)
+
+```
+### Model 22
+
+```{r}
+
+m22 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                 speaker*scale(card_sort_c, scale = F) + speaker*scale(list_sort_c, scale = F) + 
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) + 
+                 speaker*rep_acc*scale(pattern_c, scale = F) +
+                 (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m22,
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m17, m22)
+
+```
+## Final Model
+
+```{r}
+
+sjPlot::tab_model(m17,
+                  dv.labels = "Perceived Listening Effort")
+
+```

### 2024-10-15 Manuscript Analysis/data_prep.R
created data dictionary for cognitive data

@@ -2,7 +2,7 @@
 
 # Author: Micah E. Hirsch, Ph.D. (mhirsch@fsu.edu)
 
-# Data: 7/24/2024
+# Data: 10/15/2024
 
 # Purpose: To load and format cognitive data and merge it with the cleaned 
 # data from https://github.com/micah-hirsch/Listening-Effort-in-Dysarthria
@@ -15,7 +15,7 @@ library(janitor)
 
 # Setting Working Directory
 
-setwd("~/Documents/LE-Cognitive-Study/Raw Data")
+setwd("D:\\Listening Effort Study\\Raw Data\\Cognitive Data")
 
 file <- list.files()
 
@@ -41,8 +41,28 @@ cog_data <- rio::import(file) |>
 
 rm(file)
 
+# Creating Data Dictionary
+
+library(datadictionary)
+
+labels <- c(
+  subject = "Participant ID",
+  list_sort_u = "Uncorrected Standardized Working Memory Score",
+  list_sort_c = "Age-Corrected Standardized Working Memory Score",
+  flanker_u = "Uncorrected Standardized Inhibitory Control Score",
+  flanker_c = "Age-Corrected Standardized Inhbitory Control Score",
+  card_sort_u = "Uncorrected Standardized Cognitive Flexibility Score",
+  card_sort_c = "Age-Correct Standardized Cognitive Flexibility Score",
+  pattern_u = "Uncorrected Standardized Processing Speed Score",
+  pattern_c = "Age-Corrected Standardized Processing Speed Score"
+)
+
+data_dict <- create_dictionary(cog_data, var_labels = labels)
+
 # Importing demographic data
 
-setwd("~/Documents/LE-Cognitive-Study/Manuscript Analysis/Cleaned Data")
+setwd("C:\\Users\\mehirsch\\Documents\\GitHub\\LE-Cognitive-Study\\Manuscript Analysis\\Cleaned Data")
 
 rio::export(cog_data, "cleaned_cog_data.csv")
+
+rio::export(data_dict, "cleaned_cog_data_dictionary.csv")

### 2024-10-15 Manuscript Analysis/manuscript_analysis.Rmd
preliminary analyses

@@
 The data cleaning code for the pupil and perceived listening effort ratings can be found in the Github repo.
-Analyses were conducted using R version 4.3.3.
+Analyses were conducted using R version 4.4.1.
 
@@
 
-library(tidyverse)
-library(rio)
-library(glmmTMB)
-library(janitor)
-library(ggpubr)
-library(patchwork)
-library(gt)
-library(gtsummary)
-library(ggridges)
-library(performance)
-library(sjPlot)
-library(ggcorrplot)
+library(tidyverse) # install.packages("tidyverse")
+library(rio) # install.packages("rio")
+library(glmmTMB) # install.packages("glmmTMB")
+library(janitor) # install.packages("janitor")
+library(ggpubr) # install.packages("ggpubr")
+library(patchwork) # install.packages("patchwork")
+library(gt) # install.packages("gt")
+library(gtsummary) # install.packages("gtsummary")
+library(ggridges) # install.packages("ggridges")
+library(performance) # install.packages("performance")
+library(sjPlot) # install.packages("sjPlot")
+library(ggcorrplot) # install.packages("ggcorrplot")
 library(mgcv) # install.packages("mgcv")
@@
   theme(legend.position = "none")
+
+```
+
+```{r}
+
+cog |>
+  dplyr::select(ends_with("_c")) |>
+  dplyr::rename('Working Memory' = list_sort_c,
+                'Inhibitory Control' = flanker_c,
+                'Cognitive Flexibility' = card_sort_c,
+                'Processing Speed' = pattern_c) |>
+  cor(use = "pairwise.complete.obs") |>
+  ggcorrplot(
+    type = "lower",
+    outline.col = "white",
+    lab = TRUE,
+    insig = "blank",
+    legend.title = "Correlation"
+  )
 
@@
                   s(time_norm, by = condition, k = 25) +
-                  s(card_sort_c, by = condition, k = 20) +
+                  s(card_sort_c, by = condition) +
                   # Random Smooth for listener
@@
                   s(time_norm, by = condition, k = 25) +
-                  s(pattern_c, by = condition, k = 15) +
+                  s(pattern_c, by = condition) +
                   # Random Smooth for listener
@@
 m17 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
-                 speaker*scale(card_sort_c, scale = F) + speaker*scale(list_sort_c, scale = F) + 
-                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
+                 speaker*card_sort_c+ speaker*list_sort_c + 
+                 rep_acc*list_sort_c + rep_acc*card_sort_c +
                  (trial_c|subject) + (trial_c|code), data = ple)
@@
 ```
+
+```{r}
+
+ggeffects::ggpredict(m17, terms = c("list_sort_c", "speaker", "rep_acc")) |> 
+  plot()
+
+```

### 2024-10-18 Manuscript Analysis/manuscript_analysis.Rmd
reran initial analyses

@@
 
+### Distribution of Cognitive Scores
+
 The information below reports the mean age-corrected standardized scores on the subtests from the NIH Toolbox cognition battery (working memory, inhibitory control, cognitive flexibility, and processing speed).
 
-```{r}
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 # Color Palette for Figures
-my_pal <- c("#033F63", "#2CB5AE", "#28666E", "#4BAA9B", "#49E9B9", "#57A773")
+my_pal <- c("#25CED1", "#0D3B66", "#F75C03", "#3B5249", "#F1C40F", "#D90368")
 
@@
 
+ggsave("Figures\\cog_distribution.png", plot = last_plot())
+
 ```
+### Correlation Among Cognitive Subtests
 
-```{r}
+```{r, warning = F, message = F}
 
@@
 
+ggsave("Figures\\cog_correlation.png", plot = last_plot())
+
 ```
 
-```{r}
+```{r, echo = F}
 
@@
 
-rm(cog_table, demo_table, demo)
+rm(cog_table, demo)
 
@@
       color = speaker,
+      linetype = rep_acc,
       fill = speaker) +
-  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
-  geom_line() +
+  geom_line(size = .8) +
   geom_vline(xintercept = 0, size = 0.35) +
@@
   coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
-  facet_wrap("rep_acc") +
   theme_bw() +
@@
        color = "Speaker",
-       fill = "Speaker",
-       linetype = "Speaker") +
-  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS"))
+       linetype = "Repetition Accuracy") +
+  scale_color_manual(values = c(my_pal[1], my_pal[2]), labels = c("Neurotypical", "ALS"))
 
@@
 
-```{r}
+Overall, the plot is showing that as working memory increases, so does pupil dilaiton. However, there are some interesting trends in the data.
+
+```{r, message = F, warning = F}
 
@@
       y = fit,
-      color = rep_acc,
-      fill = rep_acc) +
-  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
-  geom_line() +
-  facet_wrap("speaker") +
+      color = speaker,
+      linetype = rep_acc) +
+  geom_line(size = .8) +
   theme_bw() +
@@
        y = "Pupil Dilation (Arbitrary Units)",
-       color = "Repetition Accuracy",
-       fill = "Repetition Accuracy",
+       color = "Speaker",
        linetype = "Repetition Accuracy") +
-  scale_color_manual(values = c(my_pal[1], my_pal[2])) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[2]))
+  scale_color_manual(values = c(my_pal[1], my_pal[2]))
 
@@
 
-```{r}
+Overall, there seems to be differing trends between the relationship between inhibitory control and pupil dilation depending on the speaker and phrase repetition accuracy.
+For the accurately recognized phrases from the control speaker, it appears that as inhibitory control increases, pupil dilation also increases.
+However, for the other three conditions, it seems there is a slight quadratic relationship, where listeners with lower and higher inhibitory control both have larger pupil dilation, but listeners with inhibitory scores around 70-85 (i.e., 1-2 SD below average standardized score 100), have the lowest pupil dilation.
+
+```{r, warning = F, message = F}
 
@@
        y = "Pupil Dilation (Arbitrary Units)",
-       color = "Repetition Accuracy",
-       fill = "Repetition Accuracy",
+       color = "Speaker",
        linetype = "Repetition Accuracy") +
-  scale_color_manual(values = c(my_pal[1], my_pal[2])) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[2]))
-
+  scale_color_manual(values = c(my_pal[1], my_pal[2]))
 ```
@@
 
-```{r}
+General trend is that as cognitive flexibility increases, so does pupil dilation.
+However, there is likely an effect of phrase repetition accuracy.
+Pupil dilation is higher for inaccurately recognized phrases (for both speakers, but especially the ALS speaker) for listeners with cognitive flexibility scores over 105.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+For Processing speed, there is a curvilinear relationship.
+Listeners with lower and higher processing speed scores tend to have higher pupil dilation, especially for ALS speaker (both accurately and inaccurately recognized phrases), and accurately recognized control phrases.
+
+```{r, warning = F, message = F}
 
@@
 m17 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
-                 speaker*card_sort_c+ speaker*list_sort_c + 
-                 rep_acc*list_sort_c + rep_acc*card_sort_c +
+                 speaker*scale(card_sort_c, scale = F) + speaker*scale(list_sort_c, scale = F) + 
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
                  (trial_c|subject) + (trial_c|code), data = ple)
@@
 
+The final best-fitting LME model for the PLE data is below.
+In general, both the interaction between working memory and cognitive flexibility with speaker was significant, indicating that listeners with higher working memory scores generally rated accurately-recognized ALS speaker as more effortful.
+However, listener with higher cognitive flexibility scores tend to rate accurately-recognized ALS speaker as less effortful.The three-way interaction for both of these cognitive measures with speaker and repetition accuracy was not significant.
+
+These cognitive measures also had a significant interaction with repetition accuracy.
+Listeners with higher working memory scores tend to rate inaccurately recognized phrases from the control speaker as less effortful.
+Additionally, listeners with higher cognitive flexibility scores tend to rate inaccurately-recognized phrases for the control speaker as more effortful.
+As mentioned before, the three way interactions were not significant indicating that the pattern of results for inaccurately recognized ALS phrases was similar.
+
 ```{r}
@@
 ```
-
-```{r}
-
-ggeffects::ggpredict(m17, terms = c("list_sort_c", "speaker", "rep_acc")) |> 
-  plot()
-
-```

### 2024-10-21 Manuscript Analysis/ASHA Poster Analyses.Rmd
first run through of asha analyses

---
title: "Poster Analysis"
author: "Micah Hirsch"
date: "2024-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document provides the code for the peak pupil dilation analysis for a poster presentation presented at the 2024 Annual Speech Language and Hearing Association Convention.

```{r}

# Loading in the needed packages

library(tidyverse) # install.packages("tidyverse")
library(rio) # install.packages("rio")
library(glmmTMB) # install.packages("glmmTMB")
library(janitor) # install.packages("janitor")
library(ggpubr) # install.packages("ggpubr")
library(patchwork) # install.packages("patchwork")
library(gt) # install.packages("gt")
library(gtsummary) # install.packages("gtsummary")
library(ggridges) # install.packages("ggridges")
library(performance) # install.packages("performance")
library(sjPlot) # install.packages("sjPlot")
library(ggcorrplot) # install.packages("ggcorrplot")
library(see) #install.packages("see")

# Loading in the cog data

cog <- rio::import("Cleaned Data/cleaned_cog_data.csv")

# Loading in the pupil data

pupil <- rio::import("Cleaned Data/cleaned_pupil_data_normalized.csv")

# Loading in the phrase accuracy data

phrase_acc <- rio::import("Cleaned Data/repetition_accuracy.csv") |>
  dplyr::mutate(rep_acc = ifelse(initial_response == "missing data", NA, 
                                 ifelse(target_number == correct_words_initial, "accurate", "inaccurate")),
                rep_acc_rel = ifelse(rel_response == "missing data", NA, 
                                 ifelse(target_number == correct_words_rel, "accurate", "inaccurate"))) |>
  dplyr::select(subject, trial, rep_acc)

```

# Peak Pupil Dilation

```{r}

peak_pupil_df <-pupil |>
  dplyr::group_by(subject, trial, speaker, code) |>
  dplyr::summarize(peak_pupil = max(normed_pupil)) |>
  dplyr::ungroup() |>
  dplyr::left_join(phrase_acc, by = c("subject", "trial")) |>
  dplyr::left_join(cog, by = "subject") |>
  dplyr::mutate(trial_c = trial - 6,
                speaker = factor(speaker, levels = c("Control", "ALS")),
                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate"))) |>
  dplyr::filter(!is.na(rep_acc))

```
## Visualization

```{r}

# Color Palette for Figures
my_pal <- c("#25CED1", "#0D3B66", "#F75C03", "#3B5249", "#F1C40F", "#D90368")

peak_pupil_df |>
  dplyr::select(subject, trial, speaker, peak_pupil, rep_acc, ends_with("_c")) |>
  dplyr::group_by(subject, speaker, rep_acc, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
  dplyr::summarize(peak_pupil = mean(peak_pupil)) |>
  dplyr::rename('Working Memory' = list_sort_c,
                'Inhibitory Control' = flanker_c,
                'Cognitive Flexibility' = card_sort_c,
                'Processing Speed' = pattern_c) |>
  tidyr::pivot_longer(cols = "Working Memory":"Processing Speed",
                     names_to = "cog_measure",
                     values_to = "score") |>
  ungroup() |>
  ggplot() +
  aes(x = score,
      y = peak_pupil,
      color = speaker,
      shape = speaker) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm") +
  facet_grid(rep_acc ~ cog_measure) +
  labs(x = "Score", y = "Peak Pupil Dilation") +
  theme_classic() +
  scale_color_manual(values = my_pal) +
  scale_shape_manual(values = c(1, 2)) +
  theme(aspect.ratio = 1)

```
## Models

### Fully Unconditional Model

```{r}

m0 <- glmmTMB(peak_pupil ~ 1 + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m0, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")

performance::icc(m0)
performance::check_model(m0)

```

```{r}

m1 <- glmmTMB(peak_pupil ~ 1 + (trial_c|subject), data = peak_pupil_df)
sjPlot::tab_model(m1, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m0, m1)

```

```{r}

m2 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + (trial_c|subject), data = peak_pupil_df)
sjPlot::tab_model(m2, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m1, m2)

```

```{r}

m3 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(list_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m3, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m2, m3)

```
```{r}

m3b <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m3b, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m3, m3b)

```

```{r}

m4 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(flanker_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m4, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m3b, m4)

```

```{r}

m5 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(card_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m5, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m3b, m5)

```

```{r}

m6 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(pattern_c,scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m6, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m3b, m6)

```

```{r}

m7 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                speaker*scale(list_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m7, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m3b, m7)

```

```{r}

m8 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                speaker*scale(list_sort_c, scale = F) + speaker*scale(flanker_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m8, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m7, m8)

```
```{r}

m9 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m9, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m7, m9)

```

```{r}

m10 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + speaker*scale(pattern_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m10, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m9, m10)

```

```{r}

m11 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m11, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m9, m11)

```

```{r}

m12 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(flanker_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m12, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m11, m12)

```

```{r}

m13 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m13, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m11, m13)

```

```{r}

m14 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(pattern_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m14, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m11, m14)

```

```{r}

m15 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(pattern_c, scale = F) +
                 speaker*rep_acc*scale(list_sort_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m15, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m14, m15)

```

```{r}

m16 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(pattern_c, scale = F) +
                 speaker*rep_acc*scale(flanker_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m16, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m14, m16)

```

```{r}

m17 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(pattern_c, scale = F) +
                 speaker*rep_acc*scale(card_sort_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m17, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m14, m17)

```

```{r}

m18 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(pattern_c, scale = F) +
                 speaker*rep_acc*scale(pattern_c, scale = F) +
                 (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m18, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::test_performance(m14, m18)

```

```{r}

m14b <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                  speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                  rep_acc*scale(pattern_c, scale = F) +
                  (1|subject), data = peak_pupil_df)
sjPlot::tab_model(m14b, pred.labels = "Intercept",
                  dv.labels = "Peak Pupil Dilation")
performance::check_model(m14b)

```

### 2024-10-24 Manuscript Analysis/ASHA Poster Analyses.Rmd
updated figure for peak pupil dilation

@@
 
-This document provides the code for the peak pupil dilation analysis for a poster presentation presented at the 2024 Annual Speech Language and Hearing Association Convention.
+This document provides the code for the peak pupil dilation  and perceived listening effort analyses for a poster presentation presented at the 2024 Annual Speech Language and Hearing Association Convention.
+The presentation information is below:
 
-```{r}
+-   **Title:** The Relationship Between Listener Cognition and Listening Effort During Dysarthric Speech Perception
+
+-   **Authors:** Samantha Bikulcius, Micah Hirsch, Stephanie Borrie, Yunjung Kim, Tyson Barrett, and Kaitlin Lansford
+
+-   **Location:** Seattle, WA
+
+-   **Date:** Saturday, Dec 7 10:00 AM
+
+```{r, warning = F, message = F}
 
@@
 ```
+
 ## Visualization
 
-```{r}
+```{r, warning = F, message = F}
 
@@
   geom_smooth(method = "lm") +
-  facet_grid(rep_acc ~ cog_measure) +
+  facet_grid(cog_measure ~ rep_acc) +
   labs(x = "Score", y = "Peak Pupil Dilation") +
@@
   scale_shape_manual(values = c(1, 2)) +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        panel.spacing.x = unit(.8, "lines"),
+        legend.position = "bottom")
+
+ggsave("Figures/peak_pupil_plot.png", plot = last_plot(), width = 3, height = 6, units = "in", scale = 1.2)
 
 ```
+
 ## Models
@@
 
+### Model 1
+
 ```{r}
@@
 
+### Model 2
+
 ```{r}
 
-m2 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + (trial_c|subject), data = peak_pupil_df)
+m2 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + (trial_c|subject), data = peak_pupil_df)
 sjPlot::tab_model(m2, pred.labels = "Intercept",
@@
 
+### Model 3
+
 ```{r}
 
-m3 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(list_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
+m3 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + scale(list_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
 sjPlot::tab_model(m3, pred.labels = "Intercept",
@@
 ```
+
+### Model 3b
+
 ```{r}
 
-m3b <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + (1|subject), data = peak_pupil_df)
+m3b <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + (1|subject), data = peak_pupil_df)
 sjPlot::tab_model(m3b, pred.labels = "Intercept",
@@
 
+### Model 4
+
 ```{r}
 
-m4 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(flanker_c, scale = F) + (1|subject), data = peak_pupil_df)
+m4 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + scale(flanker_c, scale = F) + (1|subject), data = peak_pupil_df)
 sjPlot::tab_model(m4, pred.labels = "Intercept",
@@
 
+### Model 5
+
 ```{r}
 
-m5 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(card_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
+m5 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + scale(card_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
 sjPlot::tab_model(m5, pred.labels = "Intercept",
@@
 
+### Model 6
+
 ```{r}
 
-m6 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + scale(pattern_c,scale = F) + (1|subject), data = peak_pupil_df)
+m6 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + scale(pattern_c,scale = F) + (1|subject), data = peak_pupil_df)
 sjPlot::tab_model(m6, pred.labels = "Intercept",
@@
 
+### Model 7
+
 ```{r}
 
-m7 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m7 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
@@
 
+### Model 8
+
 ```{r}
 
-m8 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m8 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(flanker_c, scale = F) + (1|subject), data = peak_pupil_df)
@@
 ```
+
+### Model 9
+
 ```{r}
 
-m9 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m9 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
@@
 
+### Model 10
+
 ```{r}
 
-m10 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m10 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + speaker*scale(pattern_c, scale = F) +
@@
 
+### Model 11
+
 ```{r}
 
-m11 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m11 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                  speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
@@
 
+### Model 12
+
 ```{r}
 
-m12 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m12 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                  speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
@@
 
+### Model 13
+
 ```{r}
 
-m13 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m13 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                  speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
@@
 
+### Model 14
+
 ```{r}
 
-m14 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m14 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                  speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
@@
 
+### Model 15
+
 ```{r}
 
-m15 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m15 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                  speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
@@
 
+### Model 16
+
 ```{r}
 
-m16 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m16 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                  speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
@@
 
+### Model 17
+
 ```{r}
 
-m17 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m17 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                  speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
@@
 
+### Model 18
+
 ```{r}
 
-m18 <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m18 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                  speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
@@
 
+### Best-Fitting Final Model
+
 ```{r}
 
-m14b <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+m14b <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + 
                   speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
@@
 ```
+
+# Perceived Listening Effort
+
+## Visualizations
+
+```{r}
+
+# Merge PLE data with rep acc and listener cog
+
+ple <- ple |>
+  dplyr::left_join(acc, by = c("subject", "trial")) |>
+  dplyr::filter(!is.na(rep_acc)) |>
+  dplyr::left_join(cog, by = "subject") |>
+  dplyr::filter(effort_rating > 0) |>
+  dplyr::filter(effort_rating <= 7) |>
+  dplyr::mutate(trial_c = trial - 6,
+                speaker = factor(speaker, levels = c("Control", "ALS")))
+
+ple |>
+  dplyr::select(subject, trial, speaker, code, effort_rating, rep_acc, ends_with("_c")) |>
+  dplyr::group_by(subject, speaker, rep_acc, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
+  dplyr::summarize(effort = mean(effort_rating)) |>
+  dplyr::rename('Working Memory' = list_sort_c,
+                'Inhibitory Control' = flanker_c,
+                'Cognitive Flexibility' = card_sort_c,
+                'Processing Speed' = pattern_c) |>
+  tidyr::pivot_longer(cols = "Working Memory":"Processing Speed",
+                     names_to = "cog_measure",
+                     values_to = "score") |>
+  ungroup() |>
+  ggplot() +
+  aes(x = score,
+      y = effort,
+      color = speaker,
+      shape = speaker) +
+  geom_point(alpha = 0.6) +
+  geom_smooth(method = "lm") +
+  facet_grid(rep_acc ~ cog_measure) +
+  labs(x = "Score", y = "Perceived Listening Effort Rating") +
+  theme_classic() +
+  scale_color_manual(values = my_pal) +
+  scale_shape_manual(values = c(1, 2)) +
+  theme(aspect.ratio = 1)
+
+```
+
+## Models
+
+### Fully Unconditional Model
+
+```{r, warning = F, message = F}
+
+m0 <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = ple)
+sjPlot::tab_model(m0, pred.labels = "Intercept",
+                  dv.labels = "Perceived Listening Effort")
+
+performance::icc(m0)
+
+```
+
+### Model 1
+
+In this step, we are adding trial order (trial_c) as a random slop for both the listener (subject) and phrase (code) random intercepts.
+The random intercept of trial order significantly improved model fit.
+
+```{r, warning = F, message = F}
+
+m1<- glmmTMB(effort_rating ~ 1 + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m1, pred.labels = "Intercept",
+                  dv.labels = "Perceived Listening Effort")
+performance::test_performance(m0, m1)
+
+```

### 2024-10-24 Manuscript Analysis/ASHA Poster Analyses.Rmd
ple plot

@@
 
-```{r}
+```{r, warning = F, messgae = F}
 
@@
 
-```{r}
+```{r, wrning = F, message = F}
 
@@
 performance::check_model(m14b)
+performance::test_performance(m14, m14b)
 
@@
 
+ple <- rio::import("Cleaned Data\\cleaned_ple_data.csv")
+
 ple <- ple |>
-  dplyr::left_join(acc, by = c("subject", "trial")) |>
+  dplyr::left_join(phrase_acc, by = c("subject", "trial")) |>
   dplyr::filter(!is.na(rep_acc)) |>
@@
   geom_smooth(method = "lm") +
-  facet_grid(rep_acc ~ cog_measure) +
+  facet_grid(cog_measure ~ rep_acc) +
   labs(x = "Score", y = "Perceived Listening Effort Rating") +
@@
   scale_shape_manual(values = c(1, 2)) +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1,
+        panel.spacing.x = unit(.8, "lines"),
+        legend.position = "bottom")
+
+ggsave("Figures/ple_plot.png", plot = last_plot(), width = 3, height = 6, units = "in", scale = 1.2)
 

### 2024-10-24 Manuscript Analysis/ASHA Poster Analyses.Rmd
updated peak pupil analysis

@@
 m2 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + (trial_c|subject), data = peak_pupil_df)
-sjPlot::tab_model(m2, pred.labels = "Intercept",
+sjPlot::tab_model(m2, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker * Repetition Accuracy"),
                   dv.labels = "Peak Pupil Dilation")
@@
 m3 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + scale(list_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m3, pred.labels = "Intercept",
+sjPlot::tab_model(m3, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Speaker * Repetition Accuracy"),
                   dv.labels = "Peak Pupil Dilation")
@@
 m3b <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m3b, pred.labels = "Intercept",
+sjPlot::tab_model(m3b,
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]", 
+                                  "Speaker * Repetition Accuracy"),
                   dv.labels = "Peak Pupil Dilation")
@@
 m4 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + scale(flanker_c, scale = F) + (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m4, pred.labels = "Intercept",
+sjPlot::tab_model(m4, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Inhibitory Control", "Speaker * Repetition Accuracy"),
                   dv.labels = "Peak Pupil Dilation")
@@
 m5 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + scale(card_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m5, pred.labels = "Intercept",
+sjPlot::tab_model(m5, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility", "Speaker * Repetition Accuracy"),
                   dv.labels = "Peak Pupil Dilation")
@@
 m6 <- glmmTMB(peak_pupil ~ trial_c + speaker*rep_acc + scale(pattern_c,scale = F) + (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m6, pred.labels = "Intercept",
+sjPlot::tab_model(m6, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Processing Speed", "Speaker * Repetition Accuracy"),
                   dv.labels = "Peak Pupil Dilation")
@@
                 speaker*scale(list_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m7, pred.labels = "Intercept",
+sjPlot::tab_model(m7, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]"),
                   dv.labels = "Peak Pupil Dilation")
@@
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(flanker_c, scale = F) + (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m8, pred.labels = "Intercept",
+sjPlot::tab_model(m8, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Inhibitory Control", 
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Inhibitory Control * Speaker [ALS]"),
                   dv.labels = "Peak Pupil Dilation")
@@
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m9, pred.labels = "Intercept",
+sjPlot::tab_model(m9, 
+                   pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility",
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]"),
                   dv.labels = "Peak Pupil Dilation")
@@
                  (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m10, pred.labels = "Intercept",
+sjPlot::tab_model(m10, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility", "Processing Speed",
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]", "Processing Speed * Speaker [ALS]"),
                   dv.labels = "Peak Pupil Dilation")
@@
                  (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m11, pred.labels = "Intercept",
+sjPlot::tab_model(m11, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility", 
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Accuracy [Inaccuraate]"),
                   dv.labels = "Peak Pupil Dilation")
@@
                  (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m12, pred.labels = "Intercept",
+sjPlot::tab_model(m12, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility", "Inhibitory Control",
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Accuracy [Inaccurate]",
+                                  "Inhibitory Control * Accuracy [Inaccurate]"),
                   dv.labels = "Peak Pupil Dilation")
@@
                  (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m13, pred.labels = "Intercept",
+sjPlot::tab_model(m13, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility",
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility * Accuracy [Inaccurate]"),
                   dv.labels = "Peak Pupil Dilation")
@@
                  (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m14, pred.labels = "Intercept",
+sjPlot::tab_model(m14, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility", "Processing Speed",
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Accuracy [Inaccurate]",
+                                  "Processing Speed * Accuracy [Inaccurate]"),
                   dv.labels = "Peak Pupil Dilation")
@@
                  (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m15, pred.labels = "Intercept",
+sjPlot::tab_model(m15, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility", "Processing Speed",
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Accuracy [Inaccurate]",
+                                  "Processing Speed * Accuracy [Inaccurate]",
+                                  "Working Memory * Speaker * Accuracy"),
                   dv.labels = "Peak Pupil Dilation")
@@
                  (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m16, pred.labels = "Intercept",
+sjPlot::tab_model(m16, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility", "Processing Speed", "Inhibitory Control",
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Accuracy [Inaccurate]",
+                                  "Processing Speed * Accuracy [Inaccurate]",
+                                  "Inhibitory Control * Speaker [ALS]",
+                                  "Inhibitory Control * Accuracy [Inaccurate]",
+                                  "Inhibitory Control * Speaker * Accuracy"),
                   dv.labels = "Peak Pupil Dilation")
@@
                  (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m17, pred.labels = "Intercept",
+sjPlot::tab_model(m17, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility", "Processing Speed",
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Accuracy [Inaccurate]",
+                                  "Processing Speed * Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility * Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility * Speaker * Accuracy"),
                   dv.labels = "Peak Pupil Dilation")
@@
                  (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m18, pred.labels = "Intercept",
+sjPlot::tab_model(m18, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility", "Processing Speed",
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Accuracy [Inaccurate]",
+                                  "Processing Speed * Accuracy [Inaccurate]",
+                                  "Processing Speed * Speaker [ALS]",
+                                  "Processing Speed * Speaker * Accuracy"),
                   dv.labels = "Peak Pupil Dilation")
@@
 
-### Best-Fitting Final Model
+### Removing the interaction between repetition accuracy and working memory
 
@@
                   (1|subject), data = peak_pupil_df)
-sjPlot::tab_model(m14b, pred.labels = "Intercept",
+sjPlot::tab_model(m14b, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility", "Processing Speed", 
+                                  "Speaker * Repetition Accuracy", "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]",
+                                  "Processing Speed * Repetition Accuracy [Inacccurate]"),
                   dv.labels = "Peak Pupil Dilation")
-performance::check_model(m14b)
 performance::test_performance(m14, m14b)
+
+```
+
+### Best Fitting Final Model
+
+```{r, warning = F, message = F}
+
+m14c <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+                  speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
+                  rep_acc*scale(pattern_c, scale = F) +
+                  (1|subject), data = peak_pupil_df)
+sjPlot::tab_model(m14c, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Working Memory", "Cognitive Flexibility", "Processing Speed", 
+                                  "Working Memory * Speaker [ALS]",
+                                  "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Processing Speed * Repetition Accuracy [Inacccurate]"),
+                  dv.labels = "Peak Pupil Dilation",
+                  show.std = T,
+                  file = "Tables/Peak_pupil_model.html")
+performance::check_model(m14c)
+performance::test_performance(m14b, m14c)
 

### 2024-10-24 Manuscript Analysis/ASHA Poster Analyses.Rmd
wrote in results interpretations

@@
 
+This is the best fitting model for the pupil dilation data. In regards to the cognitive data, here are the key take-aways:
+
+- Interaction between Working Memory and Speaker: On average, when listening to the ALS speaker, listeners with higher working memory scores had greater peak pupil dilation.
+- Interaction between Cognitive Flexibility and Speaker: On average, when listening to the ALS speaker, listeners with greater cognitive flexibility scores had larger peak pupil dilation values.
+- Interaction between processing speed and repetition accuracy: On average, when the phrase was inaccurately recognized, listeners with greater processing speed had greater peak pupil dilation.
+- The three way interaction between speaker, repetition accuracy, and each cognitive measure of interest was tested.
+However, none of them were significant.
+This indicates that the pattern in pupil response did not differ much between accurate and inaccurately recognized phrases for the ALS speaker (for working memory and cognitive flexibility).
+Additionally, the pattern of pupil response did not significantly differ between inaccurately-recognized control and ALS speakers for processing speed.
+
 ```{r, warning = F, message = F}

### 2024-10-24 Manuscript Analysis/ASHA Poster Analyses.Rmd
edited peak pupil dilation plot

@@
   theme_classic() +
-  scale_color_manual(values = my_pal) +
-  scale_shape_manual(values = c(1, 2)) +
+  scale_color_manual(values = my_pal, labels = c("Neurotypical", "ALS") ) +
+  scale_shape_manual(values = c(1, 2), labels = c("Neurotypical", "ALS")) +
   theme(aspect.ratio = 1,

### 2024-10-24 Manuscript Analysis/ASHA Poster Analyses.Rmd
rendered r markdown file

@@
 date: "2024-10-21"
-output: html_document
+output: 
+  html_document:
+    toc: true
+    toc_float: true
+    code_folding: hide
 ---
@@
 
-This is the best fitting model for the pupil dilation data. In regards to the cognitive data, here are the key take-aways:
+This is the best fitting model for the pupil dilation data. Model assumptions were also visually checked using the performance package. In regards to the cognitive data, here are the key take-aways:
 
@@
   theme_classic() +
-  scale_color_manual(values = my_pal) +
-  scale_shape_manual(values = c(1, 2)) +
+  scale_color_manual(values = my_pal, labels = c("Neurotypical", "ALS")) +
+  scale_shape_manual(values = c(1, 2), labels = c("Neurotypical", "ALS")) +
   theme(aspect.ratio = 1,
@@
 m0 <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = ple)
-sjPlot::tab_model(m0, pred.labels = "Intercept",
+sjPlot::tab_model(m0, 
+                  pred.labels = "Intercept",
                   dv.labels = "Perceived Listening Effort")
-
 performance::icc(m0)
+performance::check_model(m0)
 
@@
 ```
+
+### Model 2
+
+```{r, warning = F, message = F}
+
+m2 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m2,
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Speaker * Repetition Accuracy"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m1, m2)
+
+```
+
+### Model 3
+
+```{r, warning = F, message = F}
+
+m3 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + scale(list_sort_c, scale = F) +(trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m3,
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory",
+                                  "Speaker * Repetition Accuracy"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m2, m3)
+
+```
+
+### Model 4
+
+```{r, warning = F, message = F}
+
+m4 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + scale(flanker_c, scale = F) +(trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m4,
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Inhibitory Control",
+                                  "Speaker * Repetition Accuracy"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m2, m4)
+
+```
+
+### Model 5
+
+```{r, warning = F, message = F}
+
+m5 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + scale(card_sort_c, scale = F) +(trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m5,
+                  pred.labels = c("Intercept", "Trial Number", "Spekaer [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Cognitive Flexibility",
+                                  "Speaker * Repetition Accuracy"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m2, m5)
+
+```
+
+### Model 6
+
+```{r, warning = F, message = F}
+
+m6 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + scale(pattern_c, scale = F) +(trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m6,
+                  pred.labels = c("Intercept", "Trial Number", "Spekaer [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Processing Speed",
+                                  "Speaker * Repetition Accuracy"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m2, m6)
+
+```
+
+### Model 7
+
+```{r}
+
+m7 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                speaker*scale(list_sort_c, scale = F) + (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m7,
+                  pred.labels = c("Intercept", "Trial Number", "Spekaer [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory",
+                                  "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m2, m7)
+
+```
+
+### Model 8
+
+```{r, warning = F, message = F}
+
+m8 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                speaker*scale(list_sort_c, scale = F) + speaker*scale(flanker_c, scale = F) + 
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m8,
+                  pred.labels = c("Intercept", "Trial Number", "Spekaer [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Inhibitory Control",
+                                  "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]", "Inhibitory Control * Speaker [ALS]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m7, m8)
+
+```
+### Model 9
+
+```{r, warning = F, message = F}
+
+m9 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m9,
+                  pred.labels = c("Intercept", "Trial Number", "Spekaer [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
+                                  "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m7, m9)
+
+```
+
+### Model 10
+
+```{r, warning = F, message = F}
+
+m10 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
+                 speaker*scale(pattern_c, scale = F) +
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m10, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
+                                  "Processing Speed",
+                                  "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m9, m10)
+
+```
+### Model 11
+
+```{r, warning = F, message = F}
+
+m11 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
+                 rep_acc*scale(list_sort_c, scale = F) +
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m11, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
+                                  "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m9, m11)
+
+```
+
+### Model 12
+
+```{r, warning = F, message = F}
+
+m12 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(flanker_c, scale = F) +
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m12, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
+                                  "Inhibitory Control", "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Repetition Accuracy [Inaccurate]",
+                                  "Inhibitory Control * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m11, m12)
+
+```
+
+### Model 13
+
+```{r, warning = F, message = F}
+
+m13 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m13, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
+                                  "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Repetition Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m11, m13)
+
+```
+### Model 14
+
+```{r, warning = F, message = F}
+
+m14 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
+                 rep_acc*scale(pattern_c, scale = F) +
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m14, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
+                                  "Processing Speed", "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Repetition Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]",
+                                  "Processing Speed * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m13, m14)
+
+```
+
+### Model 15
+
+```{r, warning = F, message = F}
+
+m15 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
+                 speaker*rep_acc*scale(list_sort_c, scale = F) +
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m15, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility", 
+                                  "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Repetition Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]",
+                                  "Working Memory * Speaker * Accuracy"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m13, m15)
+
+```
+
+### Model 16
+
+```{r, warning = F, message = F}
+
+m16 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
+                 speaker*rep_acc*scale(flanker_c, scale = F) +
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m16, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility", 
+                                  "Inhibitory Control", "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Repetition Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]",
+                                  "Inhibitory Control * Speaker [ALS]", 
+                                  "Inhibitory Control * Repetition Accuracy [Inaccurate]",
+                                  "Inhibitory Control * Speaker * Accuracy"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m13, m16)
+
+```
+
+### Model 17
+
+```{r, warning = F, message = F}
+
+m17 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
+                 speaker*rep_acc*scale(card_sort_c, scale = F) +
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m17, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility", 
+                                  "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Repetition Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility * Speaker * Accuracy"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m13, m17)
+
+```
+
+### Model 18
+
+```{r,warning = F, message = F}
+
+m18 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
+                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
+                 speaker*rep_acc*scale(pattern_c, scale = F) +
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m18, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility", 
+                                  "Processing Speed", "Speaker * Repetition Accuracy",
+                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Repetition Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]",
+                                  "Processing Speed * Speaker [ALS]",
+                                  "Processing Speed * Repetition Accuracy [Inaccurate]",
+                                  "Processing Speed * Speaker * Accuracy"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m13, m18)
+
+```
+
+### Best Fitting Final Model
+
+Based on the model-building steps above, model 13 is the best fitting model for the perceived listening effort ratings.
+Since the interaction between speaker and repetition accuracy was not significant during any model building steps, that was removed from the model.
+Removal of this interaction did not significantly impact model performance.
+Model assumptions were also visually checked.
+Most assumptions appear to be met.
+
+The final model take-aways, in reference to the cognitve measures of interest, are:
+- The interaction between working memory and speaker: On average, when listening to the speaker with ALS, listeners with higher working memory scores rated effort higher. 
+- The interaction between cognitive flexibility and speaker: On average, when listening to the speaker with ALS, listeners with higher cognitive flexibility scores rated effort lower.
+- The interaction between working memory and repetition accuracy: On average, listeners with higher working memory scores rated effort phrases they did not accurately recognized lower.
+- The interaction between cognitive flexibility and repetition accuracy: On average, listeners with higher cognitive flexibility scores rated effort higher for phrases they did not accurately recognize.
+- Teh three way interactions between speaker, repetition accuracy, and each of the cognitve measures was tested. None of these interactions were significant.
+
+```{r, warning = F, message = F}
+
+m13b <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
+                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
+                (trial_c|subject) + (trial_c|code), data = ple)
+sjPlot::tab_model(m13b, 
+                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
+                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
+                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
+                                  "Working Memory * Repetition Accuracy [Inaccurate]",
+                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort",
+                  show.std = T,
+                  file = "Tables/ple_model.html")
+
+performance::test_performance(m13, m13b)
+performance::check_model(m0)
+
+```

### 2024-10-24 Manuscript Analysis/manuscript_analysis.Rmd
edited a few figures

@@
   scale_color_manual(values = c(my_pal[1], my_pal[2]), labels = c("Neurotypical", "ALS"))
+
+ggsave("Figures/pupil_dilation_plot.png", plot = last_plot())
 
@@
        linetype = "Repetition Accuracy") +
-  scale_color_manual(values = c(my_pal[1], my_pal[2]))
+  scale_color_manual(values = c(my_pal[1], my_pal[2]), labels = c("Neurotypical", "ALS"))
+
+ggsave("Figures/processing_speed_pupil.png", plot = last_plot())
 
@@
 
-```{r}
-
-# Merge PLE data with rep acc and listener cog
+```{r, warning = F, message = F}
 
@@
                 speaker = factor(speaker, levels = c("Control", "ALS")))
+
+ple |>
+  dplyr::group_by(speaker, rep_acc) |>
+  dplyr::summarize(per_effort = mean(effort_rating, na.rm = T), 
+                   sd = sd(effort_rating, na.rm = T),
+                   se = sd/sqrt(n())) |>
+  ggplot() +
+   aes(x = rep_acc,
+       y = per_effort,
+       group = speaker,
+       color = speaker,
+       fill = speaker) +
+  geom_bar(stat = "identity", alpha = 0.6, position = position_dodge()) +
+  geom_errorbar(aes(ymin = per_effort - se, ymax = per_effort + se), width = 0.4, position = position_dodge(.9)) +
+  labs(x = "Accuracy", y = "Perceived Listening Effort Rating", color = "Speaker", fill = "Speaker") +
+  theme_bw() +
+  theme(aspect.ratio = 1) +
+  scale_color_manual(values = c(my_pal[1], my_pal[2]), labels =  c("Neurotypical", "ALS")) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[2]), labels = c("Neurotypical", "ALS"))
+
+ggsave("Figures/ple_plot.png", plot = last_plot())
+
+```
+
+```{r, warning = F, message = F}
 

### 2024-10-28 Manuscript Analysis/ASHA Poster Analyses.Rmd
updated ple figures

@@
 
-ggsave("Figures/ple_plot.png", plot = last_plot(), width = 3, height = 6, units = "in", scale = 1.2)
+ggsave("Figures/ple_plot_cog.png", plot = last_plot(), width = 3, height = 6, units = "in", scale = 1.2)
 
@@
 ```
+
 ### Model 9
@@
 ```
+
 ### Model 11
@@
 ```
+
 ### Model 14
@@
 
-The final model take-aways, in reference to the cognitve measures of interest, are:
+The final model take-aways, in reference to the cognitive measures of interest, are:
+
 -   The interaction between working memory and speaker: On average, when listening to the speaker with ALS, listeners with higher working memory scores rated effort higher.
+
 -   The interaction between cognitive flexibility and speaker: On average, when listening to the speaker with ALS, listeners with higher cognitive flexibility scores rated effort lower.
+
 -   The interaction between working memory and repetition accuracy: On average, listeners with higher working memory scores rated effort phrases they did not accurately recognized lower.
+
 -   The interaction between cognitive flexibility and repetition accuracy: On average, listeners with higher cognitive flexibility scores rated effort higher for phrases they did not accurately recognize.
-- Teh three way interactions between speaker, repetition accuracy, and each of the cognitve measures was tested. None of these interactions were significant.
+
+-   The three way interactions between speaker, repetition accuracy, and each of the cognitive measures was tested. None of these interactions were significant.
 

### 2024-10-28 Manuscript Analysis/manuscript_analysis.Rmd

@@
   theme(aspect.ratio = 1)
+
+ggsave("Figures/ple_plot_cog.png", plot = last_plot())
 
